### 2024-03-14 paper/paper.qmd
paper qmd file created- added the data section

---
title: "Who will win the US 2024 election?!!"
subtitle: "My subtitle if needed"
author:
  - Monica Sainani
  - Kaavya Kalani
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

# Data

##Introduction to the Data:
With a focus on the American Community Survey (ACS) 2022 sample,
we present a summary of the dataset utilized in our statistical research study in this part.
Using a 1-in-100 national random sample, the ACS 2022 sample provides a thorough overview of the demographic,
socioeconomic, and geographic features of the US population. Most importantly, this dataset includes people who live in
both home and group quarters, offering a comprehensive view of the American population.

Since the ACS 2022 sample is weighted, the data are guaranteed to accurately reflect the demographic makeup of the
country's population. Weighting improves the reliability and accuracy of statistical analyses and conclusions generated
from the dataset by facilitating more precise extrapolation of findings from the sample to the larger population.

The ACS 2022 sample is noteworthy for its geographic structure, where the Public Use Microdata Area (PUMA) is the
smallest recognizable geographic unit.
Geographic areas that do not cross state boundaries and have a population of at least 100,000 are known as PUMAs.
This physical separation makes it easier to analyze regional differences and trends in socioeconomic and demographic
traits, enabling scholars to investigate the subtleties of voting patterns and electoral dynamics throughout the nation.

All things considered, the ACS 2022 sample is a useful tool for statistical analysis and empirical study, providing
insights into the complex aspects of American culture. In the sections that follow, we explore the variables that make
up the dataset and consider how important they are to comprehending voting patterns and electoral results in relation
to the 2022 presidential election.

## Introduction to the Variables
These are he key variables utilized in our statistical analysis of voting behavior during the 2022 presidential
election. These variables are drawn from the American Community Survey (ACS) 2022 sample and encompass demographic,
socio-economic, and geographic characteristics that may influence individuals' voting preferences.

Age: Age is a representation of each candidate's actual age as of the election. A key demographic factor that might
affect voting patterns, generational dynamics, and political opinions is age.

Income: The pre-tax amount of the respondents' families' entire annual income from all sources is shown as income.
As a stand-in for socioeconomic position and financial security, income has the power to influence people's political
choices, attitudes, and worldviews.

Education: Based on the highest year of school or degree earned, respondents' educational attainment is indicated by
their education. knowledge electoral dynamics requires a knowledge of the important variable of educational attainment,
which is directly connected to political ideology, civic participation, and voter turnout.

State: The respondents' state of residence provides information about where their housing units are located and makes
it possible to analyze local politics and voting trends. The political climate, the priorities of policies, and the
demographic makeup of a state can all affect how people vote and how elections turn out.

Sex: The report of sex indicates a person's gender. Political beliefs and actions are greatly influenced by gender
dynamics, which may have an impact on voting patterns and election results.

### 2024-03-14 paper/paper.qmd
paper qmd file created- added the data section

@@
 # Introduction
+A key turning point in the political environment of the United States occurred with the 2022 presidential election,
+which was marked by extreme conflict and serious economic and social challenges. We analyze data from the American
+Community Survey (ACS) 2022 sample, which offers a thorough understanding of demographic, socioeconomic, and geographic
+aspects across the nation, in order to comprehend the patterns underlying voting choices during this 2024 election
+period.
+
+Our estimand is the relationship between demographic, socio-economic, and geographic variables (such as age, income,
+education, state of residence, and gender) and voting behavior during the 2022 presidential election. Using this
+extensive dataset, our goal is to identify trends and factors that influence voting behaviour, clarifying the
+complex relationship between personal traits and local factors that influences election results.
+
+Our research adds to a greater understanding of the factors influencing voting decisions through comprehensive
+statistical analysis, yielding insightful data that may influence future election campaigns and policy discussions.
+Through highlighting the subtleties of voting patterns in the 2022 election, our research provides a critical insight
+on the condition of American democracy and the factors motivating involvement in politics and engagement.
 

### 2024-03-14 paper/paper.qmd
addded abstract

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study investigates the dynamics of voting behaviour during the US presidential election in 2022 using
+ data from the American Community Survey(ACS) 2022 sample.
+
+ We look at the correlation between voting preferences and demographic, socioeconomic, and geographic factors,
+ such as age, income, education, state of residency, and gender.
+
+ Our study reveals complex trends and determinants of voting behaviour, providing understanding into the complex
+ relationships between personal traits and local factors that influence election results.
+
+ This research contributes to a better understanding of the factors influencing political engagement in American
+ democracy by offering insightful information that might guide future electoral strategies and policy discussion."
  format: pdf

### 2024-03-14 paper/paper.qmd
model created and poststratification added

@@
 title: "Who will win the US 2024 election?!!"
-subtitle: "My subtitle if needed"
 author:
@@
 
+library(rstan)
 library(tidyverse)
-library(palmerpenguins)
+library(arrow)
+library(dplyr)
+library(knitr)
+library(ggplot2)
+
+poststrat_data <- arrow::read_parquet("data/analysis_data/poststrat_data.parquet")
+survey_data <- arrow::read_parquet("data/analysis_data/survey_data.parquet")
+us_election_model <- readRDS(file = "models/single_bay.rds")
 ```
@@
 dynamics, which may have an impact on voting patterns and election results.
+
+## Post stratification
+```{r}
+# Calculate proportions by state
+proportions_state <- poststrat_data %>%
+  group_by(state, sex, age, education, income) %>%
+  summarise(n = n()) %>%
+  group_by(state) %>%
+  mutate(prop = n/sum(n))
+
+# Obtain predictions from the Bayesian model
+proportions_state$estimate <- predict(us_election_model, newdata = proportions_state, type = "response")
+proportions_state <- proportions_state %>% mutate(num_voters = n*estimate)
+
+# Calculate prediction errors and intervals
+errors <- predict(us_election_model, newdata = proportions_state, type = "response", se.fit = TRUE)
+upper_state <- errors$fit + errors$se.fit
+lower_state <- errors$fit - errors$se.fit
+proportion_error_state <- cbind(proportions_state, lower_bound = lower_state, upper_bound = upper_state) %>%
+  rename(lower_bound = "lower_bound", upper_bound = "upper_bound")
+
+proportion_error_state <- proportion_error_state %>%
+  mutate(lower_estimate = n*lower_bound, upper_estimate = n*upper_bound)
+
+# Calculate votes for Biden
+vote_for_biden <- proportion_error_state %>%
+  mutate(biden_predict_prop = estimate * prop,
+         biden_predict_prop_lower = lower_bound * prop,
+         biden_predict_prop_upper = upper_bound * prop) %>%
+  group_by(state) %>%
+  summarise(biden_predict = sum(biden_predict_prop),
+            biden_lower = sum(biden_predict_prop_lower),
+            biden_upper = sum(biden_predict_prop_upper))
+
+# Prepare data for plotting
+state_final_split <- survey_data %>%
+  group_by(state, vote_biden) %>%
+  summarise(n = n()) %>%
+  mutate(prop = n / sum(n))
+
+state_vote_for_biden <- inner_join(state_final_split, vote_for_biden, by = "state") %>%
+  filter(vote_biden == '1')
+
+# Create the plot
+ggplot(state_vote_for_biden, aes(y = biden_predict, x = reorder(state, -biden_predict), color = "MRP estimate")) +
+  geom_point() +
+  geom_errorbar(aes(ymin = biden_lower, ymax = biden_upper), width = 0) +
+  geom_hline(yintercept = 0.5, linetype = "dashed") +
+  labs(
+    x = "State",
+    y = "Estimated proportion support for Biden",
+    color = "Source"
+  ) +
+  theme_classic() +
+  scale_color_brewer(palette = "Set1") +
+  coord_flip() +
+  theme(legend.position = "bottom")
+
+```
+
+```{r}
+# Calculate proportions by age
+proportions_age <- poststrat_data %>%
+  group_by(state, sex, age, education, income) %>%
+  summarise(n = n()) %>% 
+  group_by(age) %>%
+  mutate(prop = n/sum(n))
+
+# Obtain predictions from the Bayesian model
+proportions_age$estimate <- predict(us_election_model, newdata = proportions_age, type = "response")
+proportions_age <- proportions_age %>% mutate(num_voters = n*estimate)
+
+# Calculate prediction errors and intervals
+errors <- predict(us_election_model, newdata = proportions_age, type = "response", se.fit = TRUE)
+upper_age <- errors$fit + errors$se.fit
+lower_age <- errors$fit - errors$se.fit
+proportion_error_age <- cbind(proportions_age, lower_bound = lower_age, upper_bound = upper_age) %>% 
+  rename(lower_bound = "lower_bound", upper_bound = "upper_bound")
+
+proportion_error_age <- proportion_error_age %>%
+  mutate(lower_estimate = n*lower_bound, upper_estimate = n*upper_bound)
+
+# Calculate votes for Biden
+vote_for_biden <- proportion_error_age %>%
+  mutate(biden_predict_prop = estimate * prop,
+         biden_predict_prop_lower = lower_bound * prop,
+         biden_predict_prop_upper = upper_bound * prop) %>%
+  group_by(age) %>%
+  summarise(biden_predict = sum(biden_predict_prop),
+            biden_lower = sum(biden_predict_prop_lower),
+            biden_upper = sum(biden_predict_prop_upper))
+
+# Prepare data for plotting
+age_final_split <- survey_data %>%
+  group_by(age, vote_biden) %>%
+  summarise(n = n()) %>%
+  mutate(prop = n / sum(n))
+
+age_vote_for_biden <- inner_join(age_final_split, vote_for_biden, by = "age") %>%
+  filter(vote_biden == '1')
+
+# Plotting
+ggplot(age_vote_for_biden, aes(x = age, y = biden_predict)) +
+  geom_col(fill = "purple") +
+  labs(title = "Proportions of Votes for Biden by age group",
+       y = "Proportion of Votes",
+       x = "Age groups")
+
+```
+
+```{r}
+# sex
+# Calculate proportions by sex
+proportions_sex <- poststrat_data %>%
+  group_by(state, sex, age, education, income) %>%
+  summarise(n = n()) %>% 
+  group_by(sex) %>%
+  mutate(prop = n/sum(n))
+
+# Obtain predictions from the Bayesian model
+proportions_sex$estimate <- predict(us_election_model, newdata = proportions_sex, type = "response")
+proportions_sex <- proportions_sex %>% mutate(num_voters = n*estimate)
+
+# Calculate prediction errors and intervals
+errors <- predict(us_election_model, newdata = proportions_sex, type = "response", se.fit = TRUE)
+upper_sex <- errors$fit + errors$se.fit
+lower_sex <- errors$fit - errors$se.fit
+proportion_error_sex <- cbind(proportions_sex, lower_bound = lower_sex, upper_bound = upper_sex) %>% 
+  rename(lower_bound = "lower_bound", upper_bound = "upper_bound")
+
+proportion_error_sex <- proportion_error_sex %>%
+  mutate(lower_estimate = n*lower_bound, upper_estimate = n*upper_bound)
+
+# Calculate votes for Biden
+vote_for_biden <- proportion_error_sex %>%
+  mutate(biden_predict_prop = estimate* prop,
+         biden_predict_prop_lower = lower_bound * prop,
+         biden_predict_prop_upper = upper_bound * prop) %>%
+  group_by(sex) %>%
+  summarise(biden_predict = sum(biden_predict_prop),
+            biden_lower = sum(biden_predict_prop_lower),
+            biden_upper = sum(biden_predict_prop_upper))
+
+# Prepare data for plotting
+sex_final_split <- survey_data %>%
+  group_by(sex, vote_biden) %>%
+  summarise(n = n()) %>%
+  mutate(prop = n / sum(n))
+
+sex_vote_for_biden <- inner_join(sex_final_split, vote_for_biden, by = "sex") %>%
+  filter(vote_biden == '1')
+
+# Plotting
+ggplot(sex_vote_for_biden, aes(x = sex, y = biden_predict)) +
+  geom_col(fill = "red") +
+  labs(title = "Proportions of Votes for Biden by sex",
+       y = "Proportion of Votes",
+       x = "Sex groups")
+
+```
+
+```{r}
+# Predictions and intervals from Bayesian model
+library(rstan)
+
+# Predictions
+general_proportion_data$estimate <- as.vector(predict(us_election_model, newdata = general_proportion_data))
+
+# Prediction Intervals
+general_errors <- as.data.frame(predict(us_election_model, newdata = general_proportion_data, se.fit = TRUE))
+general_proportion_data$lower_bound <- general_errors$fit - general_errors$se.fit
+general_proportion_data$upper_bound <- general_errors$fit + general_errors$se.fit
+
+# Summarize predictions for Biden
+general_vote_for_biden <- general_proportion_data %>%
+  summarise(biden_predict = sum(estimate * prop),
+            biden_lower = sum(lower_bound * prop),
+            biden_upper = sum(upper_bound * prop))
+
+# Plotting
+general_data <- survey_data %>%
+  group_by(vote_biden) %>%
+  summarise(n = n()) %>%
+  mutate(prop = n / sum(n))
+
+ggplot(general_data, aes(x = as.factor(vote_biden), y = prop)) +
+  geom_col(fill = "blue") +
+  labs(title = "Proportions of Votes",
+       y = "Proportion",
+       x = "Vote") +
+  scale_x_discrete(labels = c("Vote for Biden", "Will not vote for Biden"))
+
+```

### 2024-03-14 scripts/03-model.R

@@ -17,7 +17,7 @@ survey_data <- arrow::read_parquet("data/analysis_data/survey_data.parquet")
 
 
 ### Model data ####
-us_election_model <-
+multi_bay <-
   stan_glmer(
     vote_biden ~ sex + (1|age) + (1|state) + (1|education) + (1|income),
     data = survey_data,
@@ -29,9 +29,42 @@ us_election_model <-
     seed = 853
   )
 
+summary(multi_bay)
+
+single_bay <-
+  stan_glm(
+    vote_biden ~ sex + age + state + education + income,
+    data = survey_data,
+    family = binomial(link = "logit"),
+    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    cores = 4,
+    adapt_delta = 0.99,
+    seed = 853
+  )
+
+summary(single_bay)
+
+freq <-  
+  glm(vote_biden ~ as.factor(sex) + as.factor(age) + as.factor(state) + as.factor(education) + as.factor(income),
+               data = survey_data,
+               family = "binomial")
+  
+summary(freq)
 #### Save model ####
 saveRDS(
-  us_election_model,
-  file = "models/us_election_model_mrp.rds"
+  multi_bay,
+  file = "models/multi_bay.rds"
+)
+
+saveRDS(
+  single_bay,
+  file = "models/single_bay.rds"
+)
+
+
+saveRDS(
+  freq,
+  file = "models/freq.rds"
 )
 

### 2024-03-14 scripts/04-poststratification.R

@@ -1,0 +1,154 @@
+#### Preamble ####
+# Purpose: Poststratification for the model
+# Author: Kaavya Kalani, Monica Sainani
+# Date: 9 March 2024
+# Contact: kaavya.kalani@mail.utoronto.ca, monica.sainani@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Run 04-model.R
+
+
+#### Workspace setup ####
+
+library(gutenbergr)
+library(haven)
+library(knitr)
+library(labelled)
+library(modelsummary)
+library(rstanarm)
+library(tidybayes)
+library(broom.mixed)
+
+poststrat_data <- arrow::read_parquet("data/analysis_data/poststrat_data.parquet")
+survey_data <- arrow::read_parquet("data/analysis_data/survey_data.parquet")
+freq <- readRDS(file = "models/single_bay.rds")
+
+# By state
+proportions_state <- poststrat_data |>
+  group_by(state, sex, age, education, income) |>
+  summarise(n = n()) |> 
+  group_by(state) |>
+  mutate(prop = n/sum(n))
+
+proportions_state$estimate <- predict.glm(freq, newdata = proportions_state, type = "response")
+proportions_state <- proportions_state |> mutate(num_voters = n*estimate)
+
+errors <- predict.glm(freq, newdata = proportions_state, type = "response", se.fit = T)
+upper_state <- errors$fit + errors$se.fit
+lower_state <- errors$fit - errors$se.fit
+proportion_error_state <- cbind(proportions_state, lower_state, upper_state) |> 
+  rename("lower_bound" = ...10, "upper_bound" = ...11)
+
+proportion_error_state <- proportion_error_state |>
+  mutate(lower_estimate = n*lower_bound, upper_estimate = n*upper_bound)
+
+vote_for_biden <- proportion_error_state |> mutate(biden_predict_prop = estimate* prop,
+                                             biden_predict_prop_lower = lower_bound * prop,
+                                             biden_predict_prop_upper = upper_bound * prop) |>
+  group_by(state) |>
+  summarise(biden_predict = sum(biden_predict_prop),
+            biden_lower = sum(biden_predict_prop_lower),
+            biden_upper = sum(biden_predict_prop_upper))
+
+state_final_split <- survey_data |>
+  group_by(state, vote_biden) |>
+  summarise(n = n()) |>
+  mutate(prop = n / sum(n))
+
[…]

### 2024-03-14 scripts/05-poststratification.R

@@ -1,63 +1,0 @@
-#### Preamble ####
-# Purpose: Poststratification for the model
-# Author: Kaavya Kalani, Monica Sainani
-# Date: 9 March 2024
-# Contact: kaavya.kalani@mail.utoronto.ca, monica.sainani@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: Run 04-model.R
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(arrow)
-
-#### Read data and model####
-poststrat_data <- arrow::read_parquet("data/analysis_data/poststrat_data.parquet")
-us_election_model <- readRDS(file = "us_election_model_mrp.rds")
-
-#### Poststratification ####
-poststrat_data_cells <-
-  poststrat_data |>
-  count(age, education, state, income, sex)
-
-poststrat_data_cells <-
-  poststrat_data_cells |>
-  mutate(prop = n / sum(n),
-         .by = state)
-
-poststrat_data_cells
-
-biden_support_by_state <-
-  us_election_model |>
-  add_epred_draws(newdata = poststrat_data_cells) |>
-  rename(support_biden_predict = .epred) |>
-  mutate(support_biden_predict_prop = support_biden_predict * prop) |>
-  ungroup() |> 
-  summarise(support_biden_predict = sum(support_biden_predict_prop),
-            .by = c(stateicp, .draw)) |>
-  summarise(
-    mean = mean(support_biden_predict),
-    lower = quantile(support_biden_predict, 0.025),
-    upper = quantile(support_biden_predict, 0.975),
-    .by = stateicp
-  )
-
-head(biden_support_by_state)
-
-#### Poststratify ####
-biden_support_by_state <-
-  us_election_model |>
-  add_epred_draws(newdata = poststrat_data_cells) |>
-  rename(support_biden_predict = .epred) |>
-  mutate(support_biden_predict_prop = support_biden_predict * prop) |>
-  ungroup() |> 
-  summarise(support_biden_predict = sum(support_biden_predict_prop),
-            .by = c(stateicp, .draw)) |>
-  summarise(
-    mean = mean(support_biden_predict),
-    lower = quantile(support_biden_predict, 0.025),
-    upper = quantile(support_biden_predict, 0.975),
-    .by = stateicp
-  )
-
-head(biden_support_by_state)

### 2024-03-14 paper/paper.qmd
more libraries added

@@
 library(ggplot2)
+library(gutenbergr)
+library(haven)
+library(labelled)
+library(modelsummary)
+library(rstanarm)
+library(tidybayes)
+library(broom.mixed)
 

### 2024-03-14 scripts/01-data_cleaning.R

@@ -7,13 +7,18 @@
 # Pre-requisites: Download the raw files as instructed in the README.md and save them in data/raw_data
 
 #### Workspace setup ####
+# install.packages("tidyverse")
+# install.packages("arrow")
 library(tidyverse)
 library(arrow)
 
 #### Clean data ####
+
+# read data to clean
 survey_data <- read_csv("data/raw_data/31120637.csv")
 poststrat_data <- arrow::read_parquet("data/raw_data/usa.parquet")
 
+# Clean and preprocess survey data
 survey_data <- survey_data %>% 
   filter(Q1_2 %in% c("Favorable", "Unfavorable")) %>% 
   mutate(age = case_when(
@@ -38,6 +43,7 @@ survey_data <- survey_data %>%
            Q1_2 == "Unfavorable" ~ 0)) %>%
   select(age, education, income, state, sex, vote_biden) 
 
+# Clean and preprocess poststratification data
 poststrat_data <- poststrat_data %>%
   mutate(AGE = as.numeric(AGE)) %>%
   filter(AGE >= 18, STATEICP != 64, STATEICP != 6, FTOTINC >= 0) %>%

### 2024-03-14 scripts/02-test_data.R

@@ -4,10 +4,12 @@
 # Date: 9 March 2024
 # Contact: kaavya.kalani@mail.utoronto.ca, monica.sainani@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Run 02-data_cleaning.R
+# Pre-requisites: Run 01-data_cleaning.R
 
 
 #### Workspace setup ####
+# install.packages("tidyverse")
+# install.packages("arrow")
 library(tidyverse)
 library(arrow)
 

### 2024-03-14 scripts/03-model.R

@@ -4,10 +4,13 @@
 # Date: 9 March 2024
 # Contact: kaavya.kalani@mail.utoronto.ca, monica.sainani@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Run 02-data_cleaning.R
+# Pre-requisites: Run 01-data_cleaning.R
 
 
 #### Workspace setup ####
+# install.packages("tidyverse")
+# install.packages("arrow")
+# install.packages("rstanarm")
 library(tidyverse)
 library(rstanarm)
 library(arrow)
@@ -17,20 +20,7 @@ survey_data <- arrow::read_parquet("data/analysis_data/survey_data.parquet")
 
 
 ### Model data ####
-multi_bay <-
-  stan_glmer(
-    vote_biden ~ sex + (1|age) + (1|state) + (1|education) + (1|income),
-    data = survey_data,
-    family = binomial(link = "logit"),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    cores = 4,
-    adapt_delta = 0.99,
-    seed = 853
-  )
-
-summary(multi_bay)
-
+# this is a single level bayesian logistic regression model
 single_bay <-
   stan_glm(
     vote_biden ~ sex + age + state + education + income,
@@ -45,26 +35,40 @@ single_bay <-
 
 summary(single_bay)
 
-freq <-  
-  glm(vote_biden ~ as.factor(sex) + as.factor(age) + as.factor(state) + as.factor(education) + as.factor(income),
-               data = survey_data,
-               family = "binomial")
-  
-summary(freq)
-#### Save model ####
-saveRDS(
-  multi_bay,
-  file = "models/multi_bay.rds"
-)
+# multi_bay <-
+#   stan_glmer(
+#     vote_biden ~ sex + (1|age) + (1|state) + (1|education) + (1|income),
+#     data = survey_data,
+#     family = binomial(link = "logit"),
+#     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+#     prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+#     cores = 4,
+#     adapt_delta = 0.99,
+#     seed = 853
+#   )
+# 
+# summary(multi_bay)
+# 
+# freq <-  
[…]

### 2024-03-14 scripts/04-poststratification.R
file path errors fixed

@@ -20,7 +20,7 @@ library(broom.mixed)
 
 poststrat_data <- arrow::read_parquet("data/analysis_data/poststrat_data.parquet")
 survey_data <- arrow::read_parquet("data/analysis_data/survey_data.parquet")
-freq <- readRDS(file = "models/single_bay.rds")
+us_election_model <- readRDS(file = "models/single_bay.rds")
 
 # By state
 proportions_state <- poststrat_data |>

### 2024-03-14 paper/paper.qmd
file path errors fixed

--- 2024-03-14 paper/paper.qmd
@@
 library(ggplot2)
+library(gutenbergr)
+library(haven)
+library(labelled)
+library(modelsummary)
+library(rstanarm)
+library(tidybayes)
+library(broom.mixed)
+library(here)
 
-poststrat_data <- arrow::read_parquet("data/analysis_data/poststrat_data.parquet")
-survey_data <- arrow::read_parquet("data/analysis_data/survey_data.parquet")
-us_election_model <- readRDS(file = "models/single_bay.rds")
+poststrat_data <- arrow::read_parquet(file = here("data/analysis_data/poststrat_data.parquet"))
+survey_data <- arrow::read_parquet(file = here("data/analysis_data/survey_data.parquet"))
+us_election_model <- readRDS(file = here("models/single_bay.rds"))
 ```
@@
 
+general_proportion_data <- poststrat_data |>
+  group_by(state, sex, age, education, income) |>
+  summarise(n = n()) |>
+  mutate(prop = n / sum(n))
+
 # Predictions

### 2024-03-14 scripts/04-poststratification.R

@@ -17,10 +17,11 @@ library(modelsummary)
 library(rstanarm)
 library(tidybayes)
 library(broom.mixed)
+library(here)
 
-poststrat_data <- arrow::read_parquet("data/analysis_data/poststrat_data.parquet")
-survey_data <- arrow::read_parquet("data/analysis_data/survey_data.parquet")
-us_election_model <- readRDS(file = "models/single_bay.rds")
+poststrat_data <- arrow::read_parquet(file = here("data/analysis_data/poststrat_data.parquet"))
+survey_data <- arrow::read_parquet(file = here("data/analysis_data/survey_data.parquet"))
+us_election_model <- readRDS(file = here("models/single_bay.rds"))
 
 # By state
 proportions_state <- poststrat_data |>

### 2024-03-14 paper/paper.qmd
format error
adding sub headings

@@
 we present a summary of the dataset utilized in our statistical research study in this part.
+
 Using a 1-in-100 national random sample, the ACS 2022 sample provides a thorough overview of the demographic,
@@
 insights into the complex aspects of American culture. In the sections that follow, we explore the variables that make
-up the dataset and consider how important they are to comprehending voting patterns and electoral results in relation
-to the 2022 presidential election.
+up the dataset and consider how important they are to comprehending voting patterns and electoral results for the 2024 election
+using this 20022 survey data.
 
@@
 ```
+
+# Results
+
+# Discussion
+
+# Appendix
+
+# References

### 2024-03-14 paper/paper.qmd
added changes to abstract and intro

@@
 date-format: long
-abstract: "This study investigates the dynamics of voting behaviour during the US presidential election in 2022 using
- data from the American Community Survey(ACS) 2022 sample.
+abstract: "This study investigates the dynamics of voting behaviour using data from the American Community Survey(ACS)
+2022 sample and predicting what the results would be for the 2024 Election.
 
  We look at the correlation between voting preferences and demographic, socioeconomic, and geographic factors,
- such as age, income, education, state of residency, and gender.
+ such as age, family income, education, state of residency, and sex.
 
@@
 # Introduction
-A key turning point in the political environment of the United States occurred with the 2022 presidential election,
+A key turning point in the political environment of the United States occurred with the 2020 presidential election,
 which was marked by extreme conflict and serious economic and social challenges. We analyze data from the American
@@
 ##Introduction to the Data:
-With a focus on the American Community Survey (ACS) 2022 sample,
-we present a summary of the dataset utilized in our statistical research study in this part.
+We will use the most recent American Community Survey, the 2022 Sample, as the post-stratification dataset
+to represent the voter population and adjust the weight of the survey data. The population for this sample was the
+entire US population, categorized into households as subgroups, and the responses of the sampled subgroups were collected.
+ACS 2022 dataset consists of 3,373,378 observations.
 
@@
 ## Introduction to the Variables
-These are he key variables utilized in our statistical analysis of voting behavior during the 2022 presidential
-election. These variables are drawn from the American Community Survey (ACS) 2022 sample and encompass demographic,
+These are the key variables utilized in our statistical analysis of voting behavior during the 2022 survey.
+These variables are drawn from the American Community Survey (ACS) 2022 sample and encompass demographic,
 socio-economic, and geographic characteristics that may influence individuals' voting preferences.
@@
 dynamics, which may have an impact on voting patterns and election results.
+
+Cleaning process:
 

### 2024-03-14 paper/paper.qmd

--- 2024-03-14 paper/paper.qmd
@@
 ## Introduction to the Variables
-These are he key variables utilized in our statistical analysis of voting behavior during the 2022 presidential
+These are the key variables utilized in our statistical analysis of voting behavior during the 2022 presidential
 election. These variables are drawn from the American Community Survey (ACS) 2022 sample and encompass demographic,
@@
 
-## Post stratification
+## Model
+We used a Bayesian Logistic Regression model with Post-Stratification to predict the 2024 US Presidential Elections.
+Logistic regression is a method used for binary classification to predict the probability of a categorical dependent variable.
+
+For our analysis, a logistic regression model will be first used to model the proportion of the voters who will vote for Biden.
+The model will be based on five independent demographic variables: state, sex, age, education and income.
+
+The logistic regression model we will be using is:
+\begin{equation}
+\log\left(\frac{\hat{p}}{1 - \hat{p}}\right) = \beta_0 + \beta_1 \times \text{state} + \beta_2 \times \text{sex} + \beta_3 \times \text{age} + \beta_4 \times \text{education} + \beta_5 \times \text{income} \label{eq:model1}
+\end{equation}
+
+where,
+
+- $\hat{p}$ represents the probability that someone will vote for Biden
+- $\beta_0$ represents the intercept term of this logistical regression. It is the probability that someone will vote for Biden if the predictors' values are zero
+- $\beta_1$ is the coefficient corresponding to state
+- $\beta_2$ is the coefficient corresponding to sex
+- $\beta_3$ is the coefficients corresponding to age
+- $\beta_4$ is the coefficients corresponding to education
+- $\beta_5$ is the coefficients corresponding to income
+
+In our model, normal priors with a mean of 0 and a standard deviation of 2.5 are used for both the coeffients and the intercept.
+Setting the mean of the priors to 0 implies that there is no expectation of a particular direction or magnitude for the coefficients or intercept.
+We chose this as we have no expectation of the same.
+The standard deviation of 2.5 reflects the uncertainty or variability in the prior beliefs.
+We chose a moderately wide prior to allow for a reasonable amount of uncertainty.
+
+The chosen priors allow the data to largely determine the posterior distribution as they are relatively non-informative.
+They don't heavily influence the results unless the data provide strong evidence to the contrary.
+
+The use of moderately wide priors can also help regularize the model, preventing overfitting and providing more stable estimates, particularly when dealing with limited data.
+
 ```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-model-coefficients
+#| tbl-cap: Coefficients of the model
+
+library(modelsummary)
+library(kableExtra)
+
+# Create model summary
+model_summary <- modelsummary(us_election_model)
+
+# Print model summary in a nice format
+print(model_summary)
+
+# Extract coefficients
+coefficients <- broom::tidy(us_election_model, conf.int = TRUE)
+coefficients |> kable(digits = 2)
+```
+
+@tbl-model-coefficients shows the coefficients for our model that can be used in the equation along with the standard error and the 95% confidence interval.
+The standard error (SE) is a measure of the precision with which a sample statistic estimates a population parameter.
+It quantifies the variability of sample statistics around the population parameter.
+A 95% confidence interval means that if you were to take many samples and construct confidence intervals in the same way, approximately 95% of those intervals would contain the true population mean.
+
+After we created the model, we then post-stratify these proportions and to conclude who we predict to win the 2024 US Presidential Elections.
+Post-stratification is a statistical technique employed to refine survey estimates by aligning them with known population characteristics.
+It helps us ensure that our findings accurately represent the diversity of the entire population.
+
+In this process, we first the organize the survey data based on the demographic factors we decided to investigate.
+This involved counting individuals in each group to understand the sample distribution and computing proportions to depict the relative contribution of each demographic group to the overall survey sample.
+
+Predictive model is then employed to generate probabilities for each demographic group, capturing the nuanced political preferences within distinct strata.
+These probabilities were summarized by aggregating them within each demographic group.
+
+Applying this post-stratification process results in enhanced predictions that align with the known population distribution within each stratum.
+These adjusted values contribute to a more accurate representation of party
+votes within specific demographic categories, ensuring our findings are reflective of the broader population.
+
+## Results
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
 # Calculate proportions by state
@@
 ```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
 # Calculate proportions by age
@@
 ```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
 # sex
@@
 ```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
 # Predictions and intervals from Bayesian model

### 2024-03-14 paper/paper.qmd
Updates

--- 2024-03-14 paper/paper.qmd
@@
 we present a summary of the dataset utilized in our statistical research study in this part.
-
 Using a 1-in-100 national random sample, the ACS 2022 sample provides a thorough overview of the demographic,
@@
 insights into the complex aspects of American culture. In the sections that follow, we explore the variables that make
-up the dataset and consider how important they are to comprehending voting patterns and electoral results for the 2024 election
-using this 20022 survey data.
+up the dataset and consider how important they are to comprehending voting patterns and electoral results in relation
+to the 2022 presidential election.
 
@@
 ```
-
-# Results
-
-# Discussion
-
-# Appendix
-
-# References

### 2024-03-15 paper/paper.qmd
Abstract updated and github link added

--- 2024-03-14 paper/paper.qmd
@@
   - Kaavya Kalani
-thanks: "Code and data are available at: LINK."
+thanks: "Code and data supporting this analysis is available at: https://github.com/sainanim/Predictions-for-the-US-2024-Elections"
 date: today
 date-format: long
-abstract: "This study investigates the dynamics of voting behaviour during the US presidential election in 2022 using
- data from the American Community Survey(ACS) 2022 sample.
-
- We look at the correlation between voting preferences and demographic, socioeconomic, and geographic factors,
- such as age, income, education, state of residency, and gender.
-
- Our study reveals complex trends and determinants of voting behaviour, providing understanding into the complex
- relationships between personal traits and local factors that influence election results.
-
- This research contributes to a better understanding of the factors influencing political engagement in American
- democracy by offering insightful information that might guide future electoral strategies and policy discussion."
+abstract: "This study investigates the dynamics of voting behaviour of americans leveraging demographic, socioeconomic, and geographic factors, such as age, family income, education, state of residency, and sex.
+We use a Bayesian Logistic Regression model with Post-Stratification to predict the outcome between Biden and Trump with the mentioned factors as predictors.
+Our study indictates that Biden is likely to be re-elected as the president with a 10% lead.
+This research contributes to a better understanding of the factors influencing political engagement in American democracy by offering insightful information that might guide future electoral strategies and policy discussion."
 format: pdf
@@
 # Introduction
-A key turning point in the political environment of the United States occurred with the 2022 presidential election,
-which was marked by extreme conflict and serious economic and social challenges. We analyze data from the American
-Community Survey (ACS) 2022 sample, which offers a thorough understanding of demographic, socioeconomic, and geographic
-aspects across the nation, in order to comprehend the patterns underlying voting choices during this 2024 election
-period.
+A key turning point in the political environment of the United States occurred with the 2020 presidential election, which was marked by extreme conflict and serious economic and social challenges.
+We analyze data from the American Community Survey (ACS) 2022 sample, which offers a thorough understanding of demographic, socioeconomic, and geographic aspects across the nation, in order to comprehend the patterns underlying voting choices during this 2024 election period.
 
-Our estimand is the relationship between demographic, socio-economic, and geographic variables (such as age, income,
-education, state of residence, and gender) and voting behavior during the 2022 presidential election. Using this
-extensive dataset, our goal is to identify trends and factors that influence voting behaviour, clarifying the
-complex relationship between personal traits and local factors that influences election results.
+Our estimand is the relationship between demographic, socio-economic, and geographic variables (such as age, income, education, state of residence, and gender) and voting behavior during the 2022 presidential election.
+Using this extensive dataset, our goal is to identify trends and factors that influence voting behaviour, clarifying the complex relationship between personal traits and local factors that influences election results.
 
@@
 statistical analysis, yielding insightful data that may influence future election campaigns and policy discussions.
-Through highlighting the subtleties of voting patterns in the 2022 election, our research provides a critical insight
-on the condition of American democracy and the factors motivating involvement in politics and engagement.
+Through highlighting the subtleties of voting patterns in the 2022 election, our research provides a critical insight on the condition of American democracy and the factors motivating involvement in politics and engagement.
 
@@
 ##Introduction to the Data:
-With a focus on the American Community Survey (ACS) 2022 sample,
-we present a summary of the dataset utilized in our statistical research study in this part.
+We will use the most recent American Community Survey, the 2022 Sample, as the post-stratification dataset
+to represent the voter population and adjust the weight of the survey data.
+The population for this sample was the entire US population, categorized into households as subgroups, and the responses of the sampled subgroups were collected.
+ACS 2022 dataset consists of 3,373,378 observations.
 
 Using a 1-in-100 national random sample, the ACS 2022 sample provides a thorough overview of the demographic,
-socioeconomic, and geographic features of the US population. Most importantly, this dataset includes people who live in
-both home and group quarters, offering a comprehensive view of the American population.
+socioeconomic, and geographic features of the US population.
+Most importantly, this dataset includes people who live in both home and group quarters, offering a comprehensive view of the American population.
 
-Since the ACS 2022 sample is weighted, the data are guaranteed to accurately reflect the demographic makeup of the
-country's population. Weighting improves the reliability and accuracy of statistical analyses and conclusions generated
-from the dataset by facilitating more precise extrapolation of findings from the sample to the larger population.
+Since the ACS 2022 sample is weighted, the data are guaranteed to accurately reflect the demographic makeup of the country's population.
+Weighting improves the reliability and accuracy of statistical analyses and conclusions generated from the dataset by facilitating more precise extrapolation of findings from the sample to the larger population.
 
@@
 Geographic areas that do not cross state boundaries and have a population of at least 100,000 are known as PUMAs.
-This physical separation makes it easier to analyze regional differences and trends in socioeconomic and demographic
-traits, enabling scholars to investigate the subtleties of voting patterns and electoral dynamics throughout the nation.
+This physical separation makes it easier to analyze regional differences and trends in socioeconomic and demographic traits, enabling scholars to investigate the subtleties of voting patterns and electoral dynamics throughout the nation.
 
-All things considered, the ACS 2022 sample is a useful tool for statistical analysis and empirical study, providing
-insights into the complex aspects of American culture. In the sections that follow, we explore the variables that make
-up the dataset and consider how important they are to comprehending voting patterns and electoral results for the 2024 election
-using this 20022 survey data.
+All things considered, the ACS 2022 sample is a useful tool for statistical analysis and empirical study, providing insights into the complex aspects of American culture.
+In the sections that follow, we explore the variables that make up the dataset and consider how important they are to comprehending voting patterns and electoral results for the 2024 election using this 20022 survey data.
 
 ## Introduction to the Variables
-These are the key variables utilized in our statistical analysis of voting behavior during the 2022 presidential
-election. These variables are drawn from the American Community Survey (ACS) 2022 sample and encompass demographic,
+These are the key variables utilized in our statistical analysis of voting behavior during the 2022 survey.
+These variables are drawn from the American Community Survey (ACS) 2022 sample and encompass demographic,
 socio-economic, and geographic characteristics that may influence individuals' voting preferences.
 
-Age: Age is a representation of each candidate's actual age as of the election. A key demographic factor that might
-affect voting patterns, generational dynamics, and political opinions is age.
+Age: Age is a representation of each candidate's actual age as of the election. A key demographic factor that might affect voting patterns, generational dynamics, and political opinions is age.
 
 Income: The pre-tax amount of the respondents' families' entire annual income from all sources is shown as income.
-As a stand-in for socioeconomic position and financial security, income has the power to influence people's political
-choices, attitudes, and worldviews.
+As a stand-in for socioeconomic position and financial security, income has the power to influence people's political choices, attitudes, and worldviews.
 
-Education: Based on the highest year of school or degree earned, respondents' educational attainment is indicated by
-their education. knowledge electoral dynamics requires a knowledge of the important variable of educational attainment,
-which is directly connected to political ideology, civic participation, and voter turnout.
+Education: Based on the highest year of school or degree earned, respondents' educational attainment is indicated by their education.
+knowledge electoral dynamics requires a knowledge of the important variable of educational attainment, which is directly connected to political ideology, civic participation, and voter turnout.
 
-State: The respondents' state of residence provides information about where their housing units are located and makes
-it possible to analyze local politics and voting trends. The political climate, the priorities of policies, and the
-demographic makeup of a state can all affect how people vote and how elections turn out.
+State: The respondents' state of residence provides information about where their housing units are located and makes it possible to analyze local politics and voting trends.
+The political climate, the priorities of policies, and the demographic makeup of a state can all affect how people vote and how elections turn out.
 
-Sex: The report of sex indicates a person's gender. Political beliefs and actions are greatly influenced by gender
-dynamics, which may have an impact on voting patterns and election results.
+Sex: The report of sex indicates a person's gender. Political beliefs and actions are greatly influenced by gender dynamics, which may have an impact on voting patterns and election results.
 

### 2024-03-15 paper/paper.qmd
Title added

@@
 ---
-title: "Who will win the US 2024 election?!!"
+title: "Biden Projected to Win in the 2024 US Presidential Election"
 author:

### 2024-03-15 paper/paper.qmd
added the discussion section

@@
 ```
+
+# Discussion
+
+## Insights Gained:
+
+Our analysis, based on sample data from the 2022 survey, followed by post-stratification focusing on key variables,
+Age, State, and Sex, yielded valuable insights into voter behavior. Notably, our findings indicated an inclination
+within the population towards voting for Biden or the Democratic Party.
+
+In the graph showcasing different regions and their voting for biden, we noticed that
+This is similar to the findings in https://www.brookings.edu/articles/forecasting-the-presidential-election-what-can-we-learn-from-the-models/
+that address the same prediction for 2024. Our model and the article predict that states that support Trump are
+from the middle and Midwest parts of the US, whereas coastal states and the Northeast tend to vote for Biden.
+Historically, rural, conservative-leaning towns with strong cultural ties to traditional values and industries like
+manufacturing and agriculture have defined the middle and Midwest regions of the United States.
+These areas frequently support gun rights, minimal government intrusion, and religious liberty—issues that are strongly
+aligned with the agenda of the Republican Party, which Trump advocated. In contrast, the Northeast and coastline states,
+which are renowned for their progressive views on culture, urban facilities, and diverse populations, have historically
+tended toward more liberal policies and social ideals that are more in line with the Democratic Party platform,
+which Biden advocated.
+
+We also created a graph highlighting different age groups and what their voting opinions are. We noticed that
+the youngest group of voters (age 18-24). The emphasis on social justice and LGBTQ+ rights among the younger generation
+may be the cause of this difference. Since there are undoubtedly more young people who identify as sexual
+minorities, there is a trend for them to support Biden and the Democratic Party. We also noticed that the older
+generation i.e individuals of age group higher than 60 years old were inclined to voting for Biden as well.
+National security and conservative principles, which are frequently connected to the Republican Party,
+may have an impact on the elder generation.
+
+Lastly, our third variable of study was Sex. We noticed a slight bias towards the democratic party from the females
+than the male gender. Women are predicted to vote for the democratic party 0.6 times more than men.
+Studies have consistently demonstrated that there is a gender difference in political ideology, with women generally
+favoring liberal or progressive policies over those of males.
+https://www.pewresearch.org/politics/2023/07/12/voting-patterns-in-the-2022-elections/#:~:text=In%20the%202022%20midterms%2C%2054,2022%20while%2051%25%20favored%20Democrats.
+This discrepancy is sometimes linked to the fact that women prioritize issues more than men do, with women giving social
+welfare, healthcare, education, and reproductive rights—all of which are more in line with the
+Democratic Party program. With its focus on women's rights, healthcare reform, and inclusivity,
+Biden's campaign may have connected more deeply with female voters, resulting in higher levels of support.
+
+## Comparative Analysis:
+Our findings align with existing literature on electoral behavior, which underscores the significance
+of demographic variables in shaping voter preferences. While our study offers insights into the inclination
+towards voting for Biden, it is important to contextualize our findings within the broader landscape of
+political science research. https://projects.economist.com/us-2020-forecast/president/how-this-works
+
+Our analysis contributes to a growing body of evidence highlighting the complex interplay between
+socio-economic factors and political attitudes, providing valuable insights for future research and
+decision-making in the political arena.
+
+## Contributions to Knowledge: (Need to edit this)
+By focusing our analysis on specific demographic variables, we contributed to a deeper understanding of the underlying
+determinants shaping electoral outcomes. Our study corroborates existing literature suggesting that demographic factors
+such as age, income, education, and geographic location play pivotal roles in influencing voter preferences.
+(https://projects.economist.com/us-2020-forecast/president/how-this-works)
+Moreover, our utilization of post-stratification techniques allowed for a more granular examination of voting behavior,
+offering insights into the heterogeneity of voter sentiment across different demographic groups.
+
+## Weaknesses and Limitations: (Need to edit this)
+
+While our analysis provides valuable insights into voter behavior, it is important to acknowledge several limitations.
+Firstly, our reliance on sample data from the 2022 survey may introduce biases and inaccuracies in our predictions.
+Our reliance on sample data from the 2022 survey may introduce a sampling bias, as the sample might not fully represent
+the diversity of the electorate. Consequently, this could lead to inaccuracies in our predictions and limit the
+generalizability of our findings to the broader population.
+
+Additionally, while we focused on key demographic variables, other influential factors such as political ideology and
+campaign messaging were not fully accounted for in our model. This means that our research has the inaccuracy of
+Incomplete Variable Consideration.
+
+Furthermore, uncertainties in the post-stratification process may impact the reliability of our estimates,
+warranting caution in interpreting the observed inclination towards voting for Biden.
+
+## Future Directions:
+Moving forward, further research is warranted to address the limitations identified in our study and advance our
+understanding of electoral dynamics. Future studies could explore the integration of additional data sources, such as
+social media sentiment analysis and polling data, to enhance the accuracy of election predictions.
+Additionally, efforts to improve transparency and reproducibility in data collection and analysis are essential for
+enhancing the reliability and validity of electoral forecasts.
+
+## Data Source:(Add where the second dataset is from)
+The data utilized in this study were sourced from the 2022 survey,
+supplemented by post-stratification data obtained from the American Community Surveys (ACS).
+These datasets provided comprehensive information on voter demographics, socio-economic characteristics,
+and geographic location, enabling a robust analysis of voting behavior in the 2020 US presidential election.

### 2024-03-15 paper/paper.qmd
sample of datasets to appendix added and intro modified

@@
 number-sections: true
+toc: true
 bibliography: references.bib
@@
 library(here)
+library(kableExtra)
 
+# read the datasets and model
 poststrat_data <- arrow::read_parquet(file = here("data/analysis_data/poststrat_data.parquet"))
@@
 Through highlighting the subtleties of voting patterns in the 2022 election, our research provides a critical insight on the condition of American democracy and the factors motivating involvement in politics and engagement.
+
+The paper is further organised into three sections: Data, Model, Results, and Discussion.
+In the Data section, we discuss how the datasets to be used for the analysis were obtained and pre-processed.
+We will explain the variables of interest in the datasets for the analysis.
+The Model section describes the model being used for the analyis.
+The Results section will then highlight and discuss the trends and associations found during the analysis.
+Lastly, the Discussion section will talk about some interesting trends found in Results in depth, link it to the real world and also highlight the weaknesses of our analysis.
 
@@
 
-library(modelsummary)
-library(kableExtra)
-
 # Create model summary
@@
 #| message: false
+#| label: fig-support-state
+#| fig-cap: Estimated proportion support for Biden by State
 
@@
 ```
+@fig-support-state
 
@@
 #| message: false
+#| label: fig-support-age
+#| fig-cap: Estimated proportion support for Biden by Age Groups
 
@@
 ```
+@fig-support-age
 
@@
 #| message: false
+#| label: fig-support-sex
+#| fig-cap: Estimated proportion support for Biden by Sex
 
@@
 ```
+@fig-support-sex
 
@@
 #| message: false
-
-# Predictions and intervals from Bayesian model
-library(rstan)
+#| label: fig-support-general
+#| fig-cap: Estimated proportion support for Biden
 
@@
 
+@fig-support-general
+
 # Discussion
@@
 
-Our analysis, based on sample data from the 2022 survey, followed by post-stratification focusing on key variables,
-Age, State, and Sex, yielded valuable insights into voter behavior. Notably, our findings indicated an inclination
-within the population towards voting for Biden or the Democratic Party.
+Our analysis, based on sample data from the 2022 survey, followed by post-stratification focusing on key variables, Age, State, and Sex, yielded valuable insights into voter behavior.
+Notably, our findings indicated an inclination within the population towards voting for Biden or the Democratic Party.
 
@@
 from the middle and Midwest parts of the US, whereas coastal states and the Northeast tend to vote for Biden.
-Historically, rural, conservative-leaning towns with strong cultural ties to traditional values and industries like
-manufacturing and agriculture have defined the middle and Midwest regions of the United States.
-These areas frequently support gun rights, minimal government intrusion, and religious liberty—issues that are strongly
-aligned with the agenda of the Republican Party, which Trump advocated. In contrast, the Northeast and coastline states,
-which are renowned for their progressive views on culture, urban facilities, and diverse populations, have historically
-tended toward more liberal policies and social ideals that are more in line with the Democratic Party platform,
-which Biden advocated.
+Historically, rural, conservative-leaning towns with strong cultural ties to traditional values and industries like manufacturing and agriculture have defined the middle and Midwest regions of the United States.
+These areas frequently support gun rights, minimal government intrusion, and religious liberty—issues that are strongly aligned with the agenda of the Republican Party, which Trump advocated.
+In contrast, the Northeast and coastline states, which are renowned for their progressive views on culture, urban facilities, and diverse populations, have historically tended toward more liberal policies and social ideals that are more in line with the Democratic Party platform, which Biden advocated.
 
 We also created a graph highlighting different age groups and what their voting opinions are. We noticed that
-the youngest group of voters (age 18-24). The emphasis on social justice and LGBTQ+ rights among the younger generation
-may be the cause of this difference. Since there are undoubtedly more young people who identify as sexual
-minorities, there is a trend for them to support Biden and the Democratic Party. We also noticed that the older
-generation i.e individuals of age group higher than 60 years old were inclined to voting for Biden as well.
-National security and conservative principles, which are frequently connected to the Republican Party,
-may have an impact on the elder generation.
+the youngest group of voters (age 18-24).
+The emphasis on social justice and LGBTQ+ rights among the younger generation may be the cause of this difference.
+Since there are undoubtedly more young people who identify as sexual minorities, there is a trend for them to support Biden and the Democratic Party.
+We also noticed that the older generation i.e individuals of age group higher than 60 years old were inclined to voting for Biden as well.
+National security and conservative principles, which are frequently connected to the Republican Party, may have an impact on the elder generation.
 
-Lastly, our third variable of study was Sex. We noticed a slight bias towards the democratic party from the females
-than the male gender. Women are predicted to vote for the democratic party 0.6 times more than men.
-Studies have consistently demonstrated that there is a gender difference in political ideology, with women generally
-favoring liberal or progressive policies over those of males.
+Lastly, our third variable of study was Sex.
+We noticed a slight bias towards the democratic party from the females than the male gender.
+Women are predicted to vote for the democratic party 0.6 times more than men.
+Studies have consistently demonstrated that there is a gender difference in political ideology, with women generally favoring liberal or progressive policies over those of males.
 https://www.pewresearch.org/politics/2023/07/12/voting-patterns-in-the-2022-elections/#:~:text=In%20the%202022%20midterms%2C%2054,2022%20while%2051%25%20favored%20Democrats.
-This discrepancy is sometimes linked to the fact that women prioritize issues more than men do, with women giving social
-welfare, healthcare, education, and reproductive rights—all of which are more in line with the
-Democratic Party program. With its focus on women's rights, healthcare reform, and inclusivity,
-Biden's campaign may have connected more deeply with female voters, resulting in higher levels of support.
+This discrepancy is sometimes linked to the fact that women prioritize issues more than men do, with women giving social welfare, healthcare, education, and reproductive rights—all of which are more in line with the Democratic Party program.
+With its focus on women's rights, healthcare reform, and inclusivity, Biden's campaign may have connected more deeply with female voters, resulting in higher levels of support.
 
 ## Comparative Analysis:
-Our findings align with existing literature on electoral behavior, which underscores the significance
-of demographic variables in shaping voter preferences. While our study offers insights into the inclination
-towards voting for Biden, it is important to contextualize our findings within the broader landscape of
-political science research. https://projects.economist.com/us-2020-forecast/president/how-this-works
+Our findings align with existing literature on electoral behavior, which underscores the significance of demographic variables in shaping voter preferences.
+While our study offers insights into the inclination towards voting for Biden, it is important to contextualize our findings within the broader landscape of political science research.
+https://projects.economist.com/us-2020-forecast/president/how-this-works
 
-Our analysis contributes to a growing body of evidence highlighting the complex interplay between
-socio-economic factors and political attitudes, providing valuable insights for future research and
-decision-making in the political arena.
+Our analysis contributes to a growing body of evidence highlighting the complex interplay between socio-economic factors and political attitudes, providing valuable insights for future research and decision-making in the political arena.
 
 ## Contributions to Knowledge: (Need to edit this)
-By focusing our analysis on specific demographic variables, we contributed to a deeper understanding of the underlying
-determinants shaping electoral outcomes. Our study corroborates existing literature suggesting that demographic factors
-such as age, income, education, and geographic location play pivotal roles in influencing voter preferences.
+By focusing our analysis on specific demographic variables, we contributed to a deeper understanding of the underlying determinants shaping electoral outcomes.
+Our study corroborates existing literature suggesting that demographic factors such as age, income, education, and geographic location play pivotal roles in influencing voter preferences.
 (https://projects.economist.com/us-2020-forecast/president/how-this-works)
-Moreover, our utilization of post-stratification techniques allowed for a more granular examination of voting behavior,
-offering insights into the heterogeneity of voter sentiment across different demographic groups.
+Moreover, our utilization of post-stratification techniques allowed for a more granular examination of voting behavior, offering insights into the heterogeneity of voter sentiment across different demographic groups.
 
@@
 Firstly, our reliance on sample data from the 2022 survey may introduce biases and inaccuracies in our predictions.
-Our reliance on sample data from the 2022 survey may introduce a sampling bias, as the sample might not fully represent
-the diversity of the electorate. Consequently, this could lead to inaccuracies in our predictions and limit the
-generalizability of our findings to the broader population.
+Our reliance on sample data from the 2022 survey may introduce a sampling bias, as the sample might not fully represent the diversity of the electorate.
+Consequently, this could lead to inaccuracies in our predictions and limit the generalizability of our findings to the broader population.
 
-Additionally, while we focused on key demographic variables, other influential factors such as political ideology and
-campaign messaging were not fully accounted for in our model. This means that our research has the inaccuracy of
-Incomplete Variable Consideration.
+Additionally, while we focused on key demographic variables, other influential factors such as political ideology and campaign messaging were not fully accounted for in our model.
+This means that our research has the inaccuracy of Incomplete Variable Consideration.
 
@@
 Moving forward, further research is warranted to address the limitations identified in our study and advance our
-understanding of electoral dynamics. Future studies could explore the integration of additional data sources, such as
-social media sentiment analysis and polling data, to enhance the accuracy of election predictions.
-Additionally, efforts to improve transparency and reproducibility in data collection and analysis are essential for
-enhancing the reliability and validity of electoral forecasts.
+understanding of electoral dynamics.
+Future studies could explore the integration of additional data sources, such as social media sentiment analysis and polling data, to enhance the accuracy of election predictions.
+Additionally, efforts to improve transparency and reproducibility in data collection and analysis are essential for enhancing the reliability and validity of electoral forecasts.
 
@@
 and geographic location, enabling a robust analysis of voting behavior in the 2020 US presidential election.
+
+# Appendix
+
+## Cleaning
+
+## Survey dataset
+Here is a glimpse of the survey data set used
+
+```{r}
+head(survey_data)
+```
+
+## Poststratification dataset
+Here is a glimpse of the poststratification data set used
+```{r}
+head(poststrat_data)
+```
+
+# References

### 2024-03-15 paper/paper.qmd
Poststratification split

@@
 us_election_model <- readRDS(file = here("models/single_bay.rds"))
+
+state_vote_for_biden <- read.csv(here("output", "state_vote_for_biden.csv"))
+
+age_vote_for_biden <- read.csv(here("output", "age_vote_for_biden.csv"))
+
+sex_vote_for_biden <- read.csv(here("output", "sex_vote_for_biden.csv"))
+
+general_data <- read.csv(here("output", "general_vote_for_biden.csv"))
+
 ```
@@
 
-# Calculate proportions by state
-proportions_state <- poststrat_data %>%
-  group_by(state, sex, age, education, income) %>%
-  summarise(n = n()) %>%
-  group_by(state) %>%
-  mutate(prop = n/sum(n))
-
-# Obtain predictions from the Bayesian model
-proportions_state$estimate <- predict(us_election_model, newdata = proportions_state, type = "response")
-proportions_state <- proportions_state %>% mutate(num_voters = n*estimate)
-
-# Calculate prediction errors and intervals
-errors <- predict(us_election_model, newdata = proportions_state, type = "response", se.fit = TRUE)
-upper_state <- errors$fit + errors$se.fit
-lower_state <- errors$fit - errors$se.fit
-proportion_error_state <- cbind(proportions_state, lower_bound = lower_state, upper_bound = upper_state) %>%
-  rename(lower_bound = "lower_bound", upper_bound = "upper_bound")
-
-proportion_error_state <- proportion_error_state %>%
-  mutate(lower_estimate = n*lower_bound, upper_estimate = n*upper_bound)
-
-# Calculate votes for Biden
-vote_for_biden <- proportion_error_state %>%
-  mutate(biden_predict_prop = estimate * prop,
-         biden_predict_prop_lower = lower_bound * prop,
-         biden_predict_prop_upper = upper_bound * prop) %>%
-  group_by(state) %>%
-  summarise(biden_predict = sum(biden_predict_prop),
-            biden_lower = sum(biden_predict_prop_lower),
-            biden_upper = sum(biden_predict_prop_upper))
-
-# Prepare data for plotting
-state_final_split <- survey_data %>%
-  group_by(state, vote_biden) %>%
-  summarise(n = n()) %>%
-  mutate(prop = n / sum(n))
-
-state_vote_for_biden <- inner_join(state_final_split, vote_for_biden, by = "state") %>%
-  filter(vote_biden == '1')
-
 # Create the plot
@@
 
-# Calculate proportions by age
-proportions_age <- poststrat_data %>%
-  group_by(state, sex, age, education, income) %>%
-  summarise(n = n()) %>% 
-  group_by(age) %>%
-  mutate(prop = n/sum(n))
-
-# Obtain predictions from the Bayesian model
-proportions_age$estimate <- predict(us_election_model, newdata = proportions_age, type = "response")
-proportions_age <- proportions_age %>% mutate(num_voters = n*estimate)
-
-# Calculate prediction errors and intervals
-errors <- predict(us_election_model, newdata = proportions_age, type = "response", se.fit = TRUE)
-upper_age <- errors$fit + errors$se.fit
-lower_age <- errors$fit - errors$se.fit
-proportion_error_age <- cbind(proportions_age, lower_bound = lower_age, upper_bound = upper_age) %>% 
-  rename(lower_bound = "lower_bound", upper_bound = "upper_bound")
-
-proportion_error_age <- proportion_error_age %>%
-  mutate(lower_estimate = n*lower_bound, upper_estimate = n*upper_bound)
-
-# Calculate votes for Biden
-vote_for_biden <- proportion_error_age %>%
-  mutate(biden_predict_prop = estimate * prop,
-         biden_predict_prop_lower = lower_bound * prop,
-         biden_predict_prop_upper = upper_bound * prop) %>%
-  group_by(age) %>%
-  summarise(biden_predict = sum(biden_predict_prop),
-            biden_lower = sum(biden_predict_prop_lower),
-            biden_upper = sum(biden_predict_prop_upper))
-
-# Prepare data for plotting
-age_final_split <- survey_data %>%
-  group_by(age, vote_biden) %>%
-  summarise(n = n()) %>%
-  mutate(prop = n / sum(n))
-
-age_vote_for_biden <- inner_join(age_final_split, vote_for_biden, by = "age") %>%
-  filter(vote_biden == '1')
-
-# Plotting
 ggplot(age_vote_for_biden, aes(x = age, y = biden_predict)) +
@@
 
-# sex
-# Calculate proportions by sex
-proportions_sex <- poststrat_data %>%
-  group_by(state, sex, age, education, income) %>%
-  summarise(n = n()) %>% 
-  group_by(sex) %>%
-  mutate(prop = n/sum(n))
-
-# Obtain predictions from the Bayesian model
-proportions_sex$estimate <- predict(us_election_model, newdata = proportions_sex, type = "response")
-proportions_sex <- proportions_sex %>% mutate(num_voters = n*estimate)
-
-# Calculate prediction errors and intervals
-errors <- predict(us_election_model, newdata = proportions_sex, type = "response", se.fit = TRUE)
-upper_sex <- errors$fit + errors$se.fit
-lower_sex <- errors$fit - errors$se.fit
-proportion_error_sex <- cbind(proportions_sex, lower_bound = lower_sex, upper_bound = upper_sex) %>% 
-  rename(lower_bound = "lower_bound", upper_bound = "upper_bound")
-
-proportion_error_sex <- proportion_error_sex %>%
-  mutate(lower_estimate = n*lower_bound, upper_estimate = n*upper_bound)
-
-# Calculate votes for Biden
-vote_for_biden <- proportion_error_sex %>%
-  mutate(biden_predict_prop = estimate* prop,
-         biden_predict_prop_lower = lower_bound * prop,
-         biden_predict_prop_upper = upper_bound * prop) %>%
-  group_by(sex) %>%
-  summarise(biden_predict = sum(biden_predict_prop),
-            biden_lower = sum(biden_predict_prop_lower),
-            biden_upper = sum(biden_predict_prop_upper))
-
-# Prepare data for plotting
-sex_final_split <- survey_data %>%
-  group_by(sex, vote_biden) %>%
-  summarise(n = n()) %>%
-  mutate(prop = n / sum(n))
-
-sex_vote_for_biden <- inner_join(sex_final_split, vote_for_biden, by = "sex") %>%
-  filter(vote_biden == '1')
-
-# Plotting
 ggplot(sex_vote_for_biden, aes(x = sex, y = biden_predict)) +
@@
 #| fig-cap: Estimated proportion support for Biden
-
-general_proportion_data <- poststrat_data |>
-  group_by(state, sex, age, education, income) |>
-  summarise(n = n()) |>
-  mutate(prop = n / sum(n))
-
-# Predictions
-general_proportion_data$estimate <- as.vector(predict(us_election_model, newdata = general_proportion_data))
-
-# Prediction Intervals
-general_errors <- as.data.frame(predict(us_election_model, newdata = general_proportion_data, se.fit = TRUE))
-general_proportion_data$lower_bound <- general_errors$fit - general_errors$se.fit
-general_proportion_data$upper_bound <- general_errors$fit + general_errors$se.fit
-
-# Summarize predictions for Biden
-general_vote_for_biden <- general_proportion_data %>%
-  summarise(biden_predict = sum(estimate * prop),
-            biden_lower = sum(lower_bound * prop),
-            biden_upper = sum(upper_bound * prop))
 

### 2024-03-15 scripts/04-poststratification.R

@@ -18,138 +18,173 @@ library(rstanarm)
 library(tidybayes)
 library(broom.mixed)
 library(here)
+library(tidyverse)
 
+# read the datasets and model
 poststrat_data <- arrow::read_parquet(file = here("data/analysis_data/poststrat_data.parquet"))
 survey_data <- arrow::read_parquet(file = here("data/analysis_data/survey_data.parquet"))
 us_election_model <- readRDS(file = here("models/single_bay.rds"))
 
-# By state
-proportions_state <- poststrat_data |>
-  group_by(state, sex, age, education, income) |>
-  summarise(n = n()) |> 
-  group_by(state) |>
+#### Post-stratification 1 (State) ####
+
+# Calculate proportions by state
+proportions_state <- poststrat_data %>%
+  group_by(state, sex, age, education, income) %>%
+  summarise(n = n()) %>%
+  group_by(state) %>%
   mutate(prop = n/sum(n))
 
-proportions_state$estimate <- predict.glm(freq, newdata = proportions_state, type = "response")
-proportions_state <- proportions_state |> mutate(num_voters = n*estimate)
+# Obtain predictions from the Bayesian model
+proportions_state$estimate <- predict(us_election_model, newdata = proportions_state, type = "response")
+proportions_state <- proportions_state %>% mutate(num_voters = n*estimate)
 
-errors <- predict.glm(freq, newdata = proportions_state, type = "response", se.fit = T)
+# Calculate prediction errors and intervals
+errors <- predict(us_election_model, newdata = proportions_state, type = "response", se.fit = TRUE)
 upper_state <- errors$fit + errors$se.fit
 lower_state <- errors$fit - errors$se.fit
-proportion_error_state <- cbind(proportions_state, lower_state, upper_state) |> 
-  rename("lower_bound" = ...10, "upper_bound" = ...11)
+proportion_error_state <- cbind(proportions_state, lower_bound = lower_state, upper_bound = upper_state) %>%
+  rename(lower_bound = "lower_bound", upper_bound = "upper_bound")
 
-proportion_error_state <- proportion_error_state |>
+proportion_error_state <- proportion_error_state %>%
   mutate(lower_estimate = n*lower_bound, upper_estimate = n*upper_bound)
 
-vote_for_biden <- proportion_error_state |> mutate(biden_predict_prop = estimate* prop,
[…]

### 2024-03-15 paper/paper.qmd
poststrat modifications

@@
 
-state_vote_for_biden <- read.csv(here("output", "state_vote_for_biden.csv"))
-
-age_vote_for_biden <- read.csv(here("output", "age_vote_for_biden.csv"))
-
-sex_vote_for_biden <- read.csv(here("output", "sex_vote_for_biden.csv"))
-
-general_data <- read.csv(here("output", "general_vote_for_biden.csv"))
+state_vote_for_biden <- read.csv(here("data/poststratified/state_vote_for_biden.csv"))
+age_vote_for_biden <- read.csv(here("data/poststratified/age_vote_for_biden.csv"))
+sex_vote_for_biden <- read.csv(here("data/poststratified/sex_vote_for_biden.csv"))
+general_data <- read.csv(here("data/poststratified/general_vote_for_biden.csv"))
 

### 2024-03-15 paper/paper.qmd
Data section added

@@
 We will explain the variables of interest in the datasets for the analysis.
-The Model section describes the model being used for the analyis.
+The Model section describes the model being used for the analysis.
 The Results section will then highlight and discuss the trends and associations found during the analysis.
@@
 
-##Introduction to the Data:
-We will use the most recent American Community Survey, the 2022 Sample, as the post-stratification dataset
-to represent the voter population and adjust the weight of the survey data.
+For this analysis, we have used two datasets.
+The datasets were cleaned and analysed using the statistical programming software R [@citeR] along with the help tidyverse [@citeTidyverse], knitr [@citeKnitr], ggplot2 [@citeGgplot], here [@citeHere], ADD MORE
+
+## Survey Data
+
+Our survey data is from the Ipsos Poll [Roper #31120637].
+It was collated from a web-based survey conducted on November 3 - 4, 2023 using KnowledgePanel.
+The survey is conducted by Ipsos and the sponsor is ABC News.
+This poll is based on a nationally representative probability sample of 949 adults age 18 or older with oversamples among Black and Hispanic respondents.
+The poll is a collection of demographic information and political preferences.
+
+A person becomes an entry in this dataset if they were 18 years or older on November 3-4, 2023, an American national and participated in this poll.
+They are recruited through a scientifically developed addressed-based sampling methodology using the latest Delivery Sequence File of the USPS – a database with full coverage of all delivery points in the US.
+Households invited to join the panel are randomly selected from all available households in the US.
+People in the sampled households are invited to join and participate in the panel.
+Those selected who do not already have internet access are provided a tablet and internet connection at no cost to the panel member.
+Those who join the panel and who are selected to participate in a survey are sent a unique password-protected log-in used to complete surveys online.
+Panelists receive a unique login to the survey and are only able to complete it one time.
+No reminder emails were sent for this study.
+
+The results from this dataset are used for our model. Among the range of variables available, we selected age, income, education, state and sex as our independent variables.
+
+- Age is the age group the panelist falls into. The poll had the ages of the panelists at the time of the poll which we then classified into one of the four age groups: 18-29, 30-44, 45-59, 60+
+
+- Income is the annual household income reported by the panelist.
+They were given the option of seven categories: Less than $10,000, $10,000 to $24,999, $25,000 to $49,999, $50,000 to $74,999, $75,000 to $99,999, $100,000 to $149,999 and $150,000 or more.
+
+- Education is the highest level of education attained by the panelist.
+They were given the option between: Less than high school, High school, Some college and Bachelor's degree or higher.
+We then classified these into three categories:  Less than high school, High school,  More than high school
+
+- State is the state of residence of the panelist.
+
+- Sex is the sex reported by the panelist and it is either male or female
+
+For the variable we are predicting, we created a binary variable which is 1 if the panelist favors Biden and 0 if the panelist favors Trump.
+There was no question in the poll which asked this direct question so we used “Overall, do you have a favorable or unfavorable impression of...Joe Biden?”.
+We acknowledge that not favoring Joe Biden does not necessarily imply favoring Donald Trump, but since they are the two major potential candidates, we classify the panelists who chose “Unfavorable” as supporting Trump, i.e not supporting Biden.
+
+## Post-stratification Data
+
+Our post-stratification data is from the most recent American Community Survey (ACS), the 2022 Sample.
+We use this to represent the voter population and adjust the weight of the survey data.
 The population for this sample was the entire US population, categorized into households as subgroups, and the responses of the sampled subgroups were collected.
 ACS 2022 dataset consists of 3,373,378 observations.
+Using a 1-in-100 national random sample, the ACS 2022 sample provides a thorough overview of the demographic, socioeconomic, and geographic features of the US population.
+Most importantly, this dataset includes people who live in both home and group quarters, offering a comprehensive view of the American population.
+A person becomes an entry in this dataset if they are a part of the sampled subgroups and have responded to the survey.
 
-Using a 1-in-100 national random sample, the ACS 2022 sample provides a thorough overview of the demographic,
-socioeconomic, and geographic features of the US population.
-Most importantly, this dataset includes people who live in both home and group quarters, offering a comprehensive view of the American population.
-
-Since the ACS 2022 sample is weighted, the data are guaranteed to accurately reflect the demographic makeup of the country's population.
-Weighting improves the reliability and accuracy of statistical analyses and conclusions generated from the dataset by facilitating more precise extrapolation of findings from the sample to the larger population.
-
-The ACS 2022 sample is noteworthy for its geographic structure, where the Public Use Microdata Area (PUMA) is the
-smallest recognizable geographic unit.
-Geographic areas that do not cross state boundaries and have a population of at least 100,000 are known as PUMAs.
-This physical separation makes it easier to analyze regional differences and trends in socioeconomic and demographic traits, enabling scholars to investigate the subtleties of voting patterns and electoral dynamics throughout the nation.
-
-All things considered, the ACS 2022 sample is a useful tool for statistical analysis and empirical study, providing insights into the complex aspects of American culture.
-In the sections that follow, we explore the variables that make up the dataset and consider how important they are to comprehending voting patterns and electoral results for the 2024 election using this 20022 survey data.
-
-## Introduction to the Variables
-These are the key variables utilized in our statistical analysis of voting behavior during the 2022 survey.
-These variables are drawn from the American Community Survey (ACS) 2022 sample and encompass demographic,
-socio-economic, and geographic characteristics that may influence individuals' voting preferences.
-
-Age: Age is a representation of each candidate's actual age as of the election. A key demographic factor that might affect voting patterns, generational dynamics, and political opinions is age.
-
-Income: The pre-tax amount of the respondents' families' entire annual income from all sources is shown as income.
-As a stand-in for socioeconomic position and financial security, income has the power to influence people's political choices, attitudes, and worldviews.
-
-Education: Based on the highest year of school or degree earned, respondents' educational attainment is indicated by their education.
-knowledge electoral dynamics requires a knowledge of the important variable of educational attainment, which is directly connected to political ideology, civic participation, and voter turnout.
-
-State: The respondents' state of residence provides information about where their housing units are located and makes it possible to analyze local politics and voting trends.
-The political climate, the priorities of policies, and the demographic makeup of a state can all affect how people vote and how elections turn out.
-
-Sex: The report of sex indicates a person's gender. Political beliefs and actions are greatly influenced by gender dynamics, which may have an impact on voting patterns and election results.
+There are a large number of different features available and the variables we chose are age, income, education, state and sex to match with the variables from our survey.
+These variables were cleaned and the dataset was preprocessed so the values and categories match with the survey dataset.
 

### 2024-03-15 paper/paper.qmd
added the discussion section

--- 2024-03-15 paper/paper.qmd
@@
 # Introduction
-A key turning point in the political environment of the United States occurred with the 2020 presidential election, which was marked by extreme conflict and serious economic and social challenges.
-We analyze data from the American Community Survey (ACS) 2022 sample, which offers a thorough understanding of demographic, socioeconomic, and geographic aspects across the nation, in order to comprehend the patterns underlying voting choices during this 2024 election period.
+A key turning point in the political environment of the United States occurred with the 2020 presidential election,
+which was marked by extreme conflict and serious economic and social challenges. We analyze data from the American
+Community Survey (ACS) 2022 sample, which offers a thorough understanding of demographic, socioeconomic, and geographic
+aspects across the nation, in order to comprehend the patterns underlying voting choices during this 2024 election period.
 
-Our estimand is the relationship between demographic, socio-economic, and geographic variables (such as age, income, education, state of residence, and gender) and voting behavior during the 2022 presidential election.
-Using this extensive dataset, our goal is to identify trends and factors that influence voting behaviour, clarifying the complex relationship between personal traits and local factors that influences election results.
+Our estimand is the relationship between demographic, socio-economic, and geographic variables
+(such as age, income, education, state of residence, and gender) and voting behavior during the 2022 presidential
+election. Using this extensive dataset, our goal is to identify trends and factors that influence voting behaviour,
+clarifying the complex relationship between personal traits and local factors that influences election results.
 
@@
 statistical analysis, yielding insightful data that may influence future election campaigns and policy discussions.
-Through highlighting the subtleties of voting patterns in the 2022 election, our research provides a critical insight on the condition of American democracy and the factors motivating involvement in politics and engagement.
+Through highlighting the subtleties of voting patterns in the 2022 election, our research provides a critical
+insight on the condition of American democracy and the factors motivating involvement in politics and engagement.
 
-The paper is further organised into three sections: Data, Model, Results, and Discussion.
-In the Data section, we discuss how the datasets to be used for the analysis were obtained and pre-processed.
+The paper is further organised into four sections: Data, Model, Results, and Discussion. In the Data section,
+ discuss how the datasets to be used for the analysis were obtained and pre-processed.
 We will explain the variables of interest in the datasets for the analysis.
@@
 
-##Introduction to the Data:
+## Introduction to the Data:
+
 We will use the most recent American Community Survey, the 2022 Sample, as the post-stratification dataset
-to represent the voter population and adjust the weight of the survey data.
-The population for this sample was the entire US population, categorized into households as subgroups, and the responses of the sampled subgroups were collected.
+to represent the voter population and adjust the weight of the survey data. The population for this sample was the
+entire US population, categorized into households as subgroups, and the responses of the sampled subgroups were collected.
 ACS 2022 dataset consists of 3,373,378 observations.
@@
 Using a 1-in-100 national random sample, the ACS 2022 sample provides a thorough overview of the demographic,
-socioeconomic, and geographic features of the US population.
-Most importantly, this dataset includes people who live in both home and group quarters, offering a comprehensive view of the American population.
+socioeconomic, and geographic features of the US population. Most importantly, this dataset includes people who live in
+both home and group quarters, offering a comprehensive view of the American population.
 
-Since the ACS 2022 sample is weighted, the data are guaranteed to accurately reflect the demographic makeup of the country's population.
-Weighting improves the reliability and accuracy of statistical analyses and conclusions generated from the dataset by facilitating more precise extrapolation of findings from the sample to the larger population.
+Since the ACS 2022 sample is weighted, the data are guaranteed to accurately reflect the demographic makeup of the
+country's population. Weighting improves the reliability and accuracy of statistical analyses and conclusions
+generated from the dataset by facilitating more precise extrapolation of findings from the sample to the larger population.
 
@@
 Geographic areas that do not cross state boundaries and have a population of at least 100,000 are known as PUMAs.
-This physical separation makes it easier to analyze regional differences and trends in socioeconomic and demographic traits, enabling scholars to investigate the subtleties of voting patterns and electoral dynamics throughout the nation.
+This physical separation makes it easier to analyze regional differences and trends in socioeconomic and demographic
+traits, enabling scholars to investigate the subtleties of voting patterns and electoral dynamics throughout the nation.
 
-All things considered, the ACS 2022 sample is a useful tool for statistical analysis and empirical study, providing insights into the complex aspects of American culture.
-In the sections that follow, we explore the variables that make up the dataset and consider how important they are to comprehending voting patterns and electoral results for the 2024 election using this 20022 survey data.
+All things considered, the ACS 2022 sample is a useful tool for statistical analysis and empirical study, providing
+insights into the complex aspects of American culture. In the sections that follow, we explore the variables that make
+up the dataset and consider how important they are to comprehending voting patterns and electoral results for the 2024
+election using this 20022 survey data.
 
@@
 
-Age: Age is a representation of each candidate's actual age as of the election. A key demographic factor that might affect voting patterns, generational dynamics, and political opinions is age.
+Age: Age is a representation of each candidate's actual age as of the election.
+A key demographic factor that might affect voting patterns, generational dynamics,
+and political opinions is age.
 
-Income: The pre-tax amount of the respondents' families' entire annual income from all sources is shown as income.
-As a stand-in for socioeconomic position and financial security, income has the power to influence people's political choices, attitudes, and worldviews.
+Income: The pre-tax amount of the respondents' families' entire annual income
+from all sources is shown as income. As a stand-in for socioeconomic position and
+financial security, income has the power to influence people's political choices, attitudes, and worldviews.
 
-Education: Based on the highest year of school or degree earned, respondents' educational attainment is indicated by their education.
-knowledge electoral dynamics requires a knowledge of the important variable of educational attainment, which is directly connected to political ideology, civic participation, and voter turnout.
+Education: Based on the highest year of school or degree earned, respondents'
+educational attainment is indicated by their education. knowledge electoral dynamics
+requires a knowledge of the important variable of educational attainment, which is
+directly connected to political ideology, civic participation, and voter turnout.
 
-State: The respondents' state of residence provides information about where their housing units are located and makes it possible to analyze local politics and voting trends.
-The political climate, the priorities of policies, and the demographic makeup of a state can all affect how people vote and how elections turn out.
+State: The respondents' state of residence provides information about where their
+housing units are located and makes it possible to analyze local politics and voting trends.
+The political climate, the priorities of policies, and the demographic makeup of a state
+can all affect how people vote and how elections turn out.
 
-Sex: The report of sex indicates a person's gender. Political beliefs and actions are greatly influenced by gender dynamics, which may have an impact on voting patterns and election results.
+Sex: The report of sex indicates a person's gender. Political beliefs and actions
+are greatly influenced by gender dynamics, which may have an impact on voting patterns and election results.
 
 ## Model
-We used a Bayesian Logistic Regression model with Post-Stratification to predict the 2024 US Presidential Elections.
-Logistic regression is a method used for binary classification to predict the probability of a categorical dependent variable.
+We used a Bayesian Logistic Regression model with Post-Stratification to predict the 2024
+US Presidential Elections. Logistic regression is a method used for binary classification to
+predict the probability of a categorical dependent variable.
 
-For our analysis, a logistic regression model will be first used to model the proportion of the voters who will vote for Biden.
-The model will be based on five independent demographic variables: state, sex, age, education and income.
+For our analysis, a logistic regression model will be first used to model the proportion
+of the voters who will vote for Biden. The model will be based on five independent
+demographic variables: state, sex, age, education and income.
 
@@
 
-@tbl-model-coefficients shows the coefficients for our model that can be used in the equation along with the standard error and the 95% confidence interval.
-The standard error (SE) is a measure of the precision with which a sample statistic estimates a population parameter.
-It quantifies the variability of sample statistics around the population parameter.
-A 95% confidence interval means that if you were to take many samples and construct confidence intervals in the same way, approximately 95% of those intervals would contain the true population mean.
+@tbl-model-coefficients shows the coefficients for our model that can be used in the equation along with the standard
+error and the 95% confidence interval. The standard error (SE) is a measure of the precision with which a sample
+statistic estimates a population parameter. It quantifies the variability of sample statistics around the population
+parameter. A 95% confidence interval means that if you were to take many samples and construct confidence intervals in
+the same way, approximately 95% of those intervals would contain the true population mean.
 
-After we created the model, we then post-stratify these proportions and to conclude who we predict to win the 2024 US Presidential Elections.
-Post-stratification is a statistical technique employed to refine survey estimates by aligning them with known population characteristics.
-It helps us ensure that our findings accurately represent the diversity of the entire population.
+After we created the model, we then post-stratify these proportions and to conclude who we predict to win the 2024 US
+Presidential Elections. Post-stratification is a statistical technique employed to refine survey estimates by aligning
+them with known population characteristics. It helps us ensure that our findings accurately represent the diversity of
+the entire population.
 
-In this process, we first the organize the survey data based on the demographic factors we decided to investigate.
-This involved counting individuals in each group to understand the sample distribution and computing proportions to depict the relative contribution of each demographic group to the overall survey sample.
+In this process, we first the organize the survey data based on the demographic factors we decided to investigate. This
+involved counting individuals in each group to understand the sample distribution and computing proportions to depict
+the relative contribution of each demographic group to the overall survey sample.
 
-Predictive model is then employed to generate probabilities for each demographic group, capturing the nuanced political preferences within distinct strata.
-These probabilities were summarized by aggregating them within each demographic group.
+Predictive model is then employed to generate probabilities for each demographic group, capturing the nuanced political
+preferences within distinct strata. These probabilities were summarized by aggregating them within each demographic
+group.
 
-Applying this post-stratification process results in enhanced predictions that align with the known population distribution within each stratum.
-These adjusted values contribute to a more accurate representation of party
+Applying this post-stratification process results in enhanced predictions that align with the known population
+distribution within each stratum. These adjusted values contribute to a more accurate representation of party
 votes within specific demographic categories, ensuring our findings are reflective of the broader population.
@@
 
-Our analysis, based on sample data from the 2022 survey, followed by post-stratification focusing on key variables, Age, State, and Sex, yielded valuable insights into voter behavior.
-Notably, our findings indicated an inclination within the population towards voting for Biden or the Democratic Party.
+Our analysis, based on sample data from the 2022 survey, followed by post-stratification focusing on key variables,
+Age, State, and Sex, yielded valuable insights into voter behavior. Notably, our findings indicated an inclination
+within the population towards voting for Biden or the Democratic Party.
 
@@
 from the middle and Midwest parts of the US, whereas coastal states and the Northeast tend to vote for Biden.
-Historically, rural, conservative-leaning towns with strong cultural ties to traditional values and industries like manufacturing and agriculture have defined the middle and Midwest regions of the United States.
-These areas frequently support gun rights, minimal government intrusion, and religious liberty—issues that are strongly aligned with the agenda of the Republican Party, which Trump advocated.
-In contrast, the Northeast and coastline states, which are renowned for their progressive views on culture, urban facilities, and diverse populations, have historically tended toward more liberal policies and social ideals that are more in line with the Democratic Party platform, which Biden advocated.
+Historically, rural, conservative-leaning towns with strong cultural ties to traditional values and industries
+like manufacturing and agriculture have defined the middle and Midwest regions of the United States.
+These areas frequently support gun rights, minimal government intrusion, and religious liberty—issues that are
+strongly aligned with the agenda of the Republican Party, which Trump advocated. In contrast, the Northeast and
+coastline states, which are renowned for their progressive views on culture, urban facilities, and diverse populations,
+have historically tended toward more liberal policies and social ideals that are more in line with the Democratic Party
+platform, which Biden advocated.
 
 We also created a graph highlighting different age groups and what their voting opinions are. We noticed that
-the youngest group of voters (age 18-24).
-The emphasis on social justice and LGBTQ+ rights among the younger generation may be the cause of this difference.
-Since there are undoubtedly more young people who identify as sexual minorities, there is a trend for them to support Biden and the Democratic Party.
-We also noticed that the older generation i.e individuals of age group higher than 60 years old were inclined to voting for Biden as well.
-National security and conservative principles, which are frequently connected to the Republican Party, may have an impact on the elder generation.
+the youngest group of voters (age 18-24). The emphasis on social justice and LGBTQ+ rights among the younger generation
+may be the cause of this difference. Since there are undoubtedly more young people who identify as sexual minorities,
+there is a trend for them to support Biden and the Democratic Party. We also noticed that the older generation i.e
+individuals of age group higher than 60 years old were inclined to voting for Biden as well.
+National security and conservative principles, which are frequently connected to the Republican Party, may have an
+impact on the elder generation.
 
-Lastly, our third variable of study was Sex.
-We noticed a slight bias towards the democratic party from the females than the male gender.
-Women are predicted to vote for the democratic party 0.6 times more than men.
-Studies have consistently demonstrated that there is a gender difference in political ideology, with women generally favoring liberal or progressive policies over those of males.
+Lastly, our third variable of study was Sex. We noticed a slight bias towards the democratic party from the females than
+the male gender. Women are predicted to vote for the democratic party 0.6 times more than men.
+Studies have consistently demonstrated that there is a gender difference in political ideology, with women generally
+favoring liberal or progressive policies over those of males.
 https://www.pewresearch.org/politics/2023/07/12/voting-patterns-in-the-2022-elections/#:~:text=In%20the%202022%20midterms%2C%2054,2022%20while%2051%25%20favored%20Democrats.
-This discrepancy is sometimes linked to the fact that women prioritize issues more than men do, with women giving social welfare, healthcare, education, and reproductive rights—all of which are more in line with the Democratic Party program.
-With its focus on women's rights, healthcare reform, and inclusivity, Biden's campaign may have connected more deeply with female voters, resulting in higher levels of support.
+This discrepancy is sometimes linked to the fact that women prioritize issues more than men do, with women giving
+social welfare, healthcare, education, and reproductive rights—all of which are more in line with the Democratic Party program.
+With its focus on women's rights, healthcare reform, and inclusivity, Biden's campaign may have connected more deeply
+with female voters, resulting in higher levels of support.
 
 ## Comparative Analysis:
-Our findings align with existing literature on electoral behavior, which underscores the significance of demographic variables in shaping voter preferences.
-While our study offers insights into the inclination towards voting for Biden, it is important to contextualize our findings within the broader landscape of political science research.
+Our findings align with existing literature on electoral behavior, which underscores the significance of demographic
+variables in shaping voter preferences. While our study offers insights into the inclination towards voting for Biden,
+it is important to contextualize our findings within the broader landscape of political science research.
 https://projects.economist.com/us-2020-forecast/president/how-this-works
 
-Our analysis contributes to a growing body of evidence highlighting the complex interplay between socio-economic factors and political attitudes, providing valuable insights for future research and decision-making in the political arena.
+Our analysis contributes to a growing body of evidence highlighting the complex interplay between socio-economic factors
+and political attitudes, providing valuable insights for future research and decision-making in the political arena.
 
 ## Contributions to Knowledge: (Need to edit this)
-By focusing our analysis on specific demographic variables, we contributed to a deeper understanding of the underlying determinants shaping electoral outcomes.
-Our study corroborates existing literature suggesting that demographic factors such as age, income, education, and geographic location play pivotal roles in influencing voter preferences.
+By focusing our analysis on specific demographic variables, we contributed to a deeper understanding of the underlying
+determinants shaping electoral outcomes. Our study corroborates existing literature suggesting that demographic
+factors such as age, income, education, and geographic location play pivotal roles in influencing voter preferences.
 (https://projects.economist.com/us-2020-forecast/president/how-this-works)
-Moreover, our utilization of post-stratification techniques allowed for a more granular examination of voting behavior, offering insights into the heterogeneity of voter sentiment across different demographic groups.
+Moreover, our utilization of post-stratification techniques allowed for a more granular examination of
+voting behavior, offering insights into the heterogeneity of voter sentiment across different demographic groups.
 
-## Weaknesses and Limitations: (Need to edit this)
+## Weaknesses and Limitations:
 
@@
 Firstly, our reliance on sample data from the 2022 survey may introduce biases and inaccuracies in our predictions.
-Our reliance on sample data from the 2022 survey may introduce a sampling bias, as the sample might not fully represent the diversity of the electorate.
-Consequently, this could lead to inaccuracies in our predictions and limit the generalizability of our findings to the broader population.
+Our reliance on sample data from the 2022 survey may introduce a sampling bias, as the sample might not fully
+represent the diversity of the electorate. Consequently, this could lead to inaccuracies in our predictions and
+limit the generalizability of our findings to the broader population.
 
-Additionally, while we focused on key demographic variables, other influential factors such as political ideology and campaign messaging were not fully accounted for in our model.
-This means that our research has the inaccuracy of Incomplete Variable Consideration.
+Additionally, while we focused on key demographic variables, other influential factors such as political ideology and
+campaign messaging were not fully accounted for in our model. This means that our research has the inaccuracy of
+Incomplete Variable Consideration.
 

### 2024-03-15 paper/paper.qmd
added the results section

--- 2024-03-15 paper/paper.qmd
@@
 @fig-support-state
+In @fig-support-state we see the percentage of the votes for Biden/Democratic party. This graph uses the data we
+collected after the post-stratification. Here, the red dots on the graph represent the voting percentage and the red
+lines are the error bars. The black vertical line with dashes indicates the line of proportion. Here, it is the equal
+split between support for Biden and support for the opposing candidate, Trump.
+
+The states that have points to the left of the black vertical line are the states voting for Trump,
+and the states on the right are voting for Biden. We see that the majority of the votes for the democratic party
+are found from the Northeast,Midwest  and the West Coast states. For example, the difference in votes for Biden
+between Oklahoma and New Mexico is more than 50%, but states like Oregon and Maine are neutral in terms of voting
+for either parties.
 
@@
 @fig-support-age
+In @fig-support-age, we used the data to find the proportion of the voters categorized by age, who would vote for the
+Democratic party. We found a pattern that the group of youngest voters and the group of the senior citizen votes are
+more likely to vote for the Democratic Party(Biden).
+
+We see that 45% of the individuals from age group 18-29 are voting
+for biden as well as 50% of the individuals from the age group of 60+ years. Meanwhile, only 35% of people in the age
+group of 45-59 are voting for Biden. Hence, in this graph we see the trend that people of ages from 30-59 are less likely
+to vote for the democratic party.
 
@@
 @fig-support-sex
+@fig-support-sex shows the proportion of Female voters that would vote for the Democratic Party as compared to the Male
+voters. We see a 10% difference in the proportion. Here, we see that Women are 10% more likely to vote
 

### 2024-03-15 paper/paper.qmd
PDF generated

@@
 
-## Model
+# Model
 We used a Bayesian Logistic Regression model with Post-Stratification to predict the 2024 US Presidential Elections.
@@
 
-# Print model summary in a nice format
-print(model_summary)
-
 # Extract coefficients
@@
 
-## Results
+# Results
 ```{r}

### 2024-03-15 paper/paper.qmd
added the results section

@@
 ```
-@fig-support-age
+
 In @fig-support-age, we used the data to find the proportion of the voters categorized by age, who would vote for the
@@
 ```
-@fig-support-sex
+
 @fig-support-sex shows the proportion of Female voters that would vote for the Democratic Party as compared to the Male
-voters. We see a 10% difference in the proportion. Here, we see that Women are 10% more likely to vote
+voters. We see more than 10% difference in the proportion. Here, we see that Women are 10% more likely to vote for Biden.
+While only 40% of the men in US would vote for Biden, more than 50% of women would vote for the Democratic Party.
 
@@
 
-@fig-support-general
+@fig-support-general looks at the division of voters for each party. The sample here is all the individuals who took
+the survey and what the proportions are in general population of the United States. We see that more than 50% of the
+population would vote for Biden or the Democratic Party and the rest would vote for Trump. This graph clearly shows the
+conclusion we can deduce from our research that with this survey done in 2022, it is clear that the population of the
+United States is inclined towards the Democratic Party.
 
@@
 
-Our analysis, based on sample data from the 2022 survey, followed by post-stratification focusing on key variables, Age, State, and Sex, yielded valuable insights into voter behavior.
-Notably, our findings indicated an inclination within the population towards voting for Biden or the Democratic Party.
+Our analysis, based on sample data from the 2022 survey, followed by post-stratification focusing on key variables, Age,
+State, and Sex, yielded valuable insights into voter behavior. Notably, our findings indicated an inclination within the
+population towards voting for Biden or the Democratic Party.
 
 In the graph showcasing different regions and their voting for biden, we noticed that
-This is similar to the findings in https://www.brookings.edu/articles/forecasting-the-presidential-election-what-can-we-learn-from-the-models/
+This is similar to the findings in https://www.brookings.edu/articles/forecasting-the-presidential-election-what-can-
+we-learn-from-the-models/
 that address the same prediction for 2024. Our model and the article predict that states that support Trump are
 from the middle and Midwest parts of the US, whereas coastal states and the Northeast tend to vote for Biden.
-Historically, rural, conservative-leaning towns with strong cultural ties to traditional values and industries like manufacturing and agriculture have defined the middle and Midwest regions of the United States.
-These areas frequently support gun rights, minimal government intrusion, and religious liberty—issues that are strongly aligned with the agenda of the Republican Party, which Trump advocated.
-In contrast, the Northeast and coastline states, which are renowned for their progressive views on culture, urban facilities, and diverse populations, have historically tended toward more liberal policies and social ideals that are more in line with the Democratic Party platform, which Biden advocated.
+Historically, rural, conservative-leaning towns with strong cultural ties to traditional values and industries like
+manufacturing and agriculture have defined the middle and Midwest regions of the United States.
+These areas frequently support gun rights, minimal government intrusion, and religious liberty—issues that are strongly
+aligned with the agenda of the Republican Party, which Trump advocated. In contrast, the Northeast and coastline states,
+which are renowned for their progressive views on culture, urban facilities, and diverse populations, have historically tended toward more liberal policies and social ideals that are more in line with the Democratic Party platform, which Biden advocated.
 

### 2024-03-15 paper/paper.qmd
made changes to the discussion section

@@
 In the graph showcasing different regions and their voting for biden, we noticed that
-This is similar to the findings in https://www.brookings.edu/articles/forecasting-the-presidential-election-what-can-
-we-learn-from-the-models/
-that address the same prediction for 2024. Our model and the article predict that states that support Trump are
-from the middle and Midwest parts of the US, whereas coastal states and the Northeast tend to vote for Biden.
+This is similar to the findings in [@James_2016] that address the same prediction for 2024. Our model and the
+article predict that states that support Trump are from the middle and Midwest parts of the US, whereas coastal states
+and the Northeast tend to vote for Biden.
+
 Historically, rural, conservative-leaning towns with strong cultural ties to traditional values and industries like
@@
 aligned with the agenda of the Republican Party, which Trump advocated. In contrast, the Northeast and coastline states,
-which are renowned for their progressive views on culture, urban facilities, and diverse populations, have historically tended toward more liberal policies and social ideals that are more in line with the Democratic Party platform, which Biden advocated.
+which are renowned for their progressive views on culture, urban facilities, and diverse populations, have historically
+tended toward more liberal policies and social ideals that are more in line with the Democratic Party platform, which Biden advocated.
 
 We also created a graph highlighting different age groups and what their voting opinions are. We noticed that
-the youngest group of voters (age 18-24).
-The emphasis on social justice and LGBTQ+ rights among the younger generation may be the cause of this difference.
-Since there are undoubtedly more young people who identify as sexual minorities, there is a trend for them to support Biden and the Democratic Party.
-We also noticed that the older generation i.e individuals of age group higher than 60 years old were inclined to voting for Biden as well.
-National security and conservative principles, which are frequently connected to the Republican Party, may have an impact on the elder generation.
+the youngest group of voters (age 18-24). The emphasis on social justice and LGBTQ+ rights among the younger generation
+may be the cause of this difference[@Brownstein_2023]. Since there are undoubtedly more young people who identify as sexual minorities,
+there is a trend for them to support Biden and the Democratic Party. We also noticed that the older generation i.e
+individuals of age group higher than 60 years old were inclined to voting for Biden as well.
+National security and conservative principles, which are frequently connected to the Republican Party,
+may have an impact on the elder generation.
 
-Lastly, our third variable of study was Sex.
-We noticed a slight bias towards the democratic party from the females than the male gender.
-Women are predicted to vote for the democratic party 0.6 times more than men.
-Studies have consistently demonstrated that there is a gender difference in political ideology, with women generally favoring liberal or progressive policies over those of males.
-https://www.pewresearch.org/politics/2023/07/12/voting-patterns-in-the-2022-elections/#:~:text=In%20the%202022%20midterms%2C%2054,2022%20while%2051%25%20favored%20Democrats.
-This discrepancy is sometimes linked to the fact that women prioritize issues more than men do, with women giving social welfare, healthcare, education, and reproductive rights—all of which are more in line with the Democratic Party program.
-With its focus on women's rights, healthcare reform, and inclusivity, Biden's campaign may have connected more deeply with female voters, resulting in higher levels of support.
+Lastly, our third variable of study was Sex. We noticed a slight bias towards the democratic party from the females than
+the male gender. Women are predicted to vote for the democratic party 0.6 times more than men.
+Studies have consistently demonstrated that there is a gender difference in political ideology, with women generally
+favoring liberal or progressive policies over those of males.[@Igielnik_2021]
+This discrepancy is sometimes linked to the fact that women prioritize issues more than men do, with women giving social
+welfare, healthcare, education, and reproductive rights—all of which are more in line with the Democratic Party program.
+With its focus on women's rights, healthcare reform, and inclusivity, Biden's campaign may have connected more
+deeply with female voters, resulting in higher levels of support.
 
 ## Comparative Analysis:
-Our findings align with existing literature on electoral behavior, which underscores the significance of demographic variables in shaping voter preferences.
-While our study offers insights into the inclination towards voting for Biden, it is important to contextualize our findings within the broader landscape of political science research.
-https://projects.economist.com/us-2020-forecast/president/how-this-works
+Our findings align with existing literature on electoral behavior, which underscores the significance of demographic
+variables in shaping voter preferences. While our study offers insights into the inclination towards voting for Biden,
+it is important to contextualize our findings within the broader landscape of political science research.[Gelman]
 
-Our analysis contributes to a growing body of evidence highlighting the complex interplay between socio-economic factors and political attitudes, providing valuable insights for future research and decision-making in the political arena.
+Our analysis contributes to a growing body of evidence highlighting the complex interplay between socio-economic factors
+and political attitudes, providing valuable insights for future research and decision-making in the political arena.
 
-## Contributions to Knowledge: (Need to edit this)
-By focusing our analysis on specific demographic variables, we contributed to a deeper understanding of the underlying determinants shaping electoral outcomes.
-Our study corroborates existing literature suggesting that demographic factors such as age, income, education, and geographic location play pivotal roles in influencing voter preferences.
-(https://projects.economist.com/us-2020-forecast/president/how-this-works)
-Moreover, our utilization of post-stratification techniques allowed for a more granular examination of voting behavior, offering insights into the heterogeneity of voter sentiment across different demographic groups.
+## Contributions to Knowledge:
+By focusing our analysis on specific demographic variables, we contributed to a deeper understanding of the underlying
+determinants shaping electoral outcomes. Our study corroborates existing literature suggesting that demographic factors
+such as age, income, education, and geographic location play pivotal roles in influencing voter preferences. [@Gelman]
+Moreover, our utilization of post-stratification techniques allowed for a more granular examination of voting behavior,
+offering insights into the heterogeneity of voter sentiment across different demographic groups.
 
-## Weaknesses and Limitations: (Need to edit this)
+## Weaknesses and Limitations:
 
@@
 Firstly, our reliance on sample data from the 2022 survey may introduce biases and inaccuracies in our predictions.
-Our reliance on sample data from the 2022 survey may introduce a sampling bias, as the sample might not fully represent the diversity of the electorate.
-Consequently, this could lead to inaccuracies in our predictions and limit the generalizability of our findings to the broader population.
+Our reliance on sample data from the 2022 survey may introduce a sampling bias, as the sample might not fully represent
+the diversity of the electorate. Consequently, this could lead to inaccuracies in our predictions and limit the
+generalizability of our findings to the broader population.
 
-Additionally, while we focused on key demographic variables, other influential factors such as political ideology and campaign messaging were not fully accounted for in our model.
-This means that our research has the inaccuracy of Incomplete Variable Consideration.
+Additionally, while we focused on key demographic variables, other influential factors such as political ideology and
+campaign messaging were not fully accounted for in our model. This means that our research has the inaccuracy of
+Incomplete Variable Consideration[@Funakoshi_2020].
 
@@
 Moving forward, further research is warranted to address the limitations identified in our study and advance our
-understanding of electoral dynamics.
-Future studies could explore the integration of additional data sources, such as social media sentiment analysis and polling data, to enhance the accuracy of election predictions.
-Additionally, efforts to improve transparency and reproducibility in data collection and analysis are essential for enhancing the reliability and validity of electoral forecasts.
+understanding of electoral dynamics. Future studies could explore the integration of additional data sources,
+such as social media sentiment analysis and polling data, to enhance the accuracy of election predictions.
+Additionally, efforts to improve transparency and reproducibility in data collection and analysis are essential
+for enhancing the reliability and validity of electoral forecasts.
 

### 2024-03-15 paper/paper.qmd
made changes to the discussion section

@@
 
-## Data Source:(Add where the second dataset is from)
-The data utilized in this study were sourced from the 2022 survey,
-supplemented by post-stratification data obtained from the American Community Surveys (ACS).
-These datasets provided comprehensive information on voter demographics, socio-economic characteristics,
-and geographic location, enabling a robust analysis of voting behavior in the 2020 US presidential election.
-
 # Appendix

### 2024-03-15 paper/paper.qmd
made changes to the introduction section

@@
 # Introduction
-A key turning point in the political environment of the United States occurred with the 2020 presidential election, which was marked by extreme conflict and serious economic and social challenges.
-We analyze data from the American Community Survey (ACS) 2022 sample, which offers a thorough understanding of demographic, socioeconomic, and geographic aspects across the nation, in order to comprehend the patterns underlying voting choices during this 2024 election period.
+A key turning point in the political environment of the United States occurred with the 2020 presidential election,
+which was marked by extreme conflict and serious economic and social challenges. We analyze data from the American
+Community Survey (ACS) 2022 sample, which offers a thorough understanding of demographic, socioeconomic,
+and geographic aspects across the nation, in order to comprehend the patterns underlying voting choices during
+this 2024 election period. This paper focuses on the two major political parties, Democratic and Republican.
 
-Our estimand is the relationship between demographic, socio-economic, and geographic variables (such as age, income, education, state of residence, and gender) and voting behavior during the 2022 presidential election.
-Using this extensive dataset, our goal is to identify trends and factors that influence voting behaviour, clarifying the complex relationship between personal traits and local factors that influences election results.
+The rivalry between the Democratic and Republican parties, as embodied by their opposing candidates,
+Joe Biden and Donald Trump, has been an iconic feature of the modern American political scene. The Democratic Party's
+nominee is the veteran politician and former Obama vice president Joe Biden. With an emphasis on middle-class families,
+Biden's program prioritizes measures that advance racial justice, healthcare reform, climate action, and economic
+regeneration. He is an advocate of a more compassionate and inclusive style of governing, hoping to bring the
+country back together and bring decorum back into political conversation.
+
+On the other hand, the current Republican Party standard-bearer and President Donald Trump represents a more populist
+and nationalist agenda. A major focus on economic nationalism, immigration restrictions, tax cuts, and
+deregulation have been highlights of Trump's leadership. His administration prioritized themes like law and order,
+trade protectionism, and border security as part of its conservative agenda. Trump's aggressive approach to politics
+is a defining characteristic of his leadership style; it frequently divides public opinion and provokes strong emotions
+from both supporters and opponents.
+
+Our estimand is the relationship between demographic, socio-economic, and geographic variables (such as age, income,
+education, state of residence, and gender) and voting behavior during the 2022 presidential election.
+Using this extensive dataset, our goal is to identify trends and factors that influence voting behaviour,
+clarifying the complex relationship between personal traits and local factors that influences election results.
 
@@
 statistical analysis, yielding insightful data that may influence future election campaigns and policy discussions.
-Through highlighting the subtleties of voting patterns in the 2022 election, our research provides a critical insight on the condition of American democracy and the factors motivating involvement in politics and engagement.
+Through highlighting the subtleties of voting patterns in the 2022 election, our research provides a critical insight
+on the condition of American democracy and the factors motivating involvement in politics and engagement.
 
-The paper is further organised into three sections: Data, Model, Results, and Discussion.
-In the Data section, we discuss how the datasets to be used for the analysis were obtained and pre-processed.
-We will explain the variables of interest in the datasets for the analysis.
-The Model section describes the model being used for the analysis.
-The Results section will then highlight and discuss the trends and associations found during the analysis.
-Lastly, the Discussion section will talk about some interesting trends found in Results in depth, link it to the real world and also highlight the weaknesses of our analysis.
+The paper is further organised into three sections: Data, Model, Results, and Discussion. In the Data section, we
+discuss how the datasets to be used for the analysis were obtained and pre-processed. We will explain the variables of
+interest in the datasets for the analysis. The Model section describes the model being used for the analysis. The Results
+section will then highlight and discuss the trends and associations found during the analysis. Lastly, the Discussion
+section will talk about some interesting trends found in Results in depth, link it to the real world and also highlight
+the weaknesses of our analysis.
 

### 2024-03-15 paper/paper.qmd
PDF rendered

@@
 voters. We see more than 10% difference in the proportion. Here, we see that Women are 10% more likely to vote for Biden.
-While only 40% of the men in US would vote for Biden, more than 50% of women would vote for the Democratic Party.
+While only 40% of the men in US would vote for Biden, more than 50% of women would vote for the Biden.
 
@@
 the survey and what the proportions are in general population of the United States. We see that more than 50% of the
-population would vote for Biden or the Democratic Party and the rest would vote for Trump. This graph clearly shows the
+population would vote for Biden and the rest would vote for Trump. This graph clearly shows the
 conclusion we can deduce from our research that with this survey done in 2022, it is clear that the population of the

### 2024-03-15 paper/paper.qmd
made changes to the introduction section

--- 2024-03-15 paper/paper.qmd
@@
 date-format: long
-abstract: "This study investigates the dynamics of voting behaviour of americans leveraging demographic, socioeconomic, and geographic factors, such as age, family income, education, state of residency, and sex.
-We use a Bayesian Logistic Regression model with Post-Stratification to predict the outcome between Biden and Trump with the mentioned factors as predictors.
-Our study indictates that Biden is likely to be re-elected as the president with a 10% lead.
-This research contributes to a better understanding of the factors influencing political engagement in American democracy by offering insightful information that might guide future electoral strategies and policy discussion."
+abstract: "This study investigates the dynamics of voting behaviour of Americans leveraging demographic, socioeconomic,
+and geographic factors, such as age, family income, education, state of residency, and sex. We use a Bayesian Logistic
+Regression model with Post-Stratification to predict the outcome between Biden and Trump with the mentioned factors as
+predictors. Our study indicates that Biden is likely to be re-elected as the president with a 10% lead. This research
+contributes to a better understanding of the factors influencing political engagement in American democracy by offering
+insightful information that might guide future electoral strategies and policy discussion."
 format: pdf

### 2024-03-15 paper/paper.qmd
made changes to the introduction section

@@
 ---
-title: "Biden Projected to Win in the 2024 US Presidential Election"
+title: "Biden Projected to Win the 2024 US Presidential Elections"
 author:

### 2024-03-15 paper/paper.qmd
Subtitle added, abstract edited

@@
 title: "Biden Projected to Win the 2024 US Presidential Elections"
+subtitle: "Analysis using data from American Community Survey 2022 and Ipsos Poll 2023 levereging Bayesian Logistic Regression with Post-Stratification"
 author:
@@
 date-format: long
-abstract: "This study investigates the dynamics of voting behaviour of Americans leveraging demographic, socioeconomic,
-and geographic factors, such as age, family income, education, state of residency, and sex. We use a Bayesian Logistic
-Regression model with Post-Stratification to predict the outcome between Biden and Trump with the mentioned factors as
-predictors. Our study indicates that Biden is likely to be re-elected as the president with a 10% lead. This research
-contributes to a better understanding of the factors influencing political engagement in American democracy by offering
-insightful information that might guide future electoral strategies and policy discussion."
+abstract: "We analyze American voting behavior, focusing on demographic, socioeconomic, and geographic factors.
+We use a Bayesian Logistic Regression model with Post-Stratification to predict the election outcome between Biden and Trump.
+We find a strong relationship between age, family income, education, state of residency, sex and likely voting outcome.
+Our central forecast is a Biden re-election with a 10% lead.
+This research provides insights into the factors affecting political engagement in the U.S., potentially guiding future electoral strategies and policy discussions."
 format: pdf

### 2024-03-16 paper/paper.qmd
added sketches

@@
 State, and Sex, yielded valuable insights into voter behavior. Notably, our findings indicated an inclination within the
-population towards voting for Biden or the Democratic Party.
+population towards voting for Biden from the Democratic Party.
 
-In the graph showcasing different regions and their voting for biden, we noticed that
-This is similar to the findings in [@James_2016] that address the same prediction for 2024. Our model and the
+In the graph showcasing different regions and their voting for Biden, we noticed that
+this is similar to the findings in an article by statistician James E. Campbell,  [@James_2016] that address the same prediction for 2024. Our model and the
 article predict that states that support Trump are from the middle and Midwest parts of the US, whereas coastal states

### 2024-03-16 scripts/00-simulate_data.R

@@ -10,10 +10,48 @@
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(dplyr)
+
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
+# Set seed for reproducibility
+set.seed(123)
+
+# Define the number of observations
+n <- 1000
+
+# Simulate person column (serial number)
+person <- 1:n
+
+# Simulate age data (ranging from 18 to 98)
+age <- sample(18:98, n, replace = TRUE)
+
+# Simulate sex data (female or male)
+sex <- sample(c("Female", "Male"), n, replace = TRUE)
+
+# Simulate family income data (ranging from 20000 to 1000000)
+income <- runif(n, min = 20000, max = 1000000)
+
+# Simulate state data (states in the US)
+states <- c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware",
+"Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine",
+"Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada",
+"New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", "Oklahoma", "Oregon",
+"Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia",
+"Washington", "West Virginia", "Wisconsin", "Wyoming")
+state <- sample(states, n, replace = TRUE)
+
+# Simulate general vote data (0 for Trump, 1 for Biden)
+general_vote <- sample(0:1, n, replace = TRUE)
+
+# Create a dataframe
+simulated_data <- data.frame(person, age, sex, income, state, general_vote)
+
+# View the first few rows of the simulated dataset
+head(simulated_data)
+
+
+
 
 
 

### 2024-03-16 scripts/00-simulate_data.R
added simulation

@@ -32,14 +32,20 @@ sex <- sample(c("Female", "Male"), n, replace = TRUE)
 # Simulate family income data (ranging from 20000 to 1000000)
 income <- runif(n, min = 20000, max = 1000000)
 
-# Simulate state data (states in the US)
-states <- c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware",
-"Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine",
-"Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada",
-"New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", "Oklahoma", "Oregon",
-"Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia",
-"Washington", "West Virginia", "Wisconsin", "Wyoming")
-state <- sample(states, n, replace = TRUE)
+# Simulate state data using the provided codes and labels
+state_codes <- c("09", "23", "25", "33", "44", "50", "10", "34", "36", "42", "01", "18", "26", "36", "55",
+"29", "20", "22", "31", "46", "47", "27", "32", "37", "38", "51", "19", "40", "28", "30", "35", "45", "60",
+"10", "11", "12", "13", "24", "41", "48", "56", "21", "54", "50", "55", "23", "24", "44", "48", "49", "15",
+"16", "17", "05", "06", "08", "78", "97", "98", "99")
+state_labels <- c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "Rhode Island",
+"Vermont", "Delaware", "New Jersey", "New York", "Pennsylvania", "Illinois", "Indiana", "Michigan", "Ohio",
+"Wisconsin", "Iowa", "Kansas", "Minnesota", "Missouri", "Nebraska", "North Dakota", "South Dakota", "Virginia",
+"Alabama", "Arkansas", "Florida", "Georgia", "Louisiana", "Mississippi", "North Carolina", "South Carolina",
+"Texas", "Kentucky", "Maryland", "Oklahoma", "Tennessee", "West Virginia", "Arizona", "Colorado", "Idaho", "Montana",
[…]

### 2024-03-16 scripts/00-simulate_data.R
Preamble added and file cleaned up

@@ -1,42 +1,23 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Simulated the survey and poststratification datasets
+# Author: Kaavya Kalani, Monica Sainani
+# Date: 9 March 2024
+# Contact: kaavya.kalani@mail.utoronto.ca, monica.sainani@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: None
 
 #### Workspace setup ####
 library(tidyverse)
 library(dplyr)
 
-
-#### Simulate data ####
-# Set seed for reproducibility
-set.seed(123)
-
-# Define the number of observations
+#### Simulate survey dataset ####
+set.seed(323)
 n <- 1000
 
-# Simulate person column (serial number)
-person <- 1:n
-
-# Simulate age data (ranging from 18 to 98)
-age <- sample(18:98, n, replace = TRUE)
-
-# Simulate sex data (female or male)
-sex <- sample(c("Female", "Male"), n, replace = TRUE)
-
-# Simulate family income data (ranging from 20000 to 1000000)
-income <- runif(n, min = 20000, max = 1000000)
-
-# Simulate state data using the provided codes and labels
-state_codes <- c("09", "23", "25", "33", "44", "50", "10", "34", "36", "42", "01", "18", "26", "36", "55",
-"29", "20", "22", "31", "46", "47", "27", "32", "37", "38", "51", "19", "40", "28", "30", "35", "45", "60",
-"10", "11", "12", "13", "24", "41", "48", "56", "21", "54", "50", "55", "23", "24", "44", "48", "49", "15",
-"16", "17", "05", "06", "08", "78", "97", "98", "99")
+age <- sample(18:98, n, replace = TRUE) # Simulate age data (ranging from 18 to 98)
+sex <- sample(c("Female", "Male"), n, replace = TRUE) # Simulate sex data (female or male)
+income <- round(runif(n, min = 20000, max = 1000000)) # Simulate family income data (ranging from 20000 to 1000000)
+ 
[…]

### 2024-03-16 paper/paper.qmd
Preamble updates, citations for all packages used added and poststrat sim added

@@
 
-library(rstan)
+# install.packages("broom.mixed")
+# install.packages("tidyverse")
+# install.packages("arrow")
+# install.packages("dplyr")
+# install.packages("knitr")
+# install.packages("ggplot2")
+# install.packages("modelsummary")
+# install.packages("here")
+# install.packages("kableExtra")
+
+library(broom.mixed)
 library(tidyverse)
@@
 library(ggplot2)
-library(gutenbergr)
-library(haven)
-library(labelled)
 library(modelsummary)
-library(rstanarm)
-library(tidybayes)
-library(broom.mixed)
 library(here)
@@
 
-# read the datasets and model
+# Read the datasets and model
 poststrat_data <- arrow::read_parquet(file = here("data/analysis_data/poststrat_data.parquet"))
@@
 
+# Read the post-stratified data
 state_vote_for_biden <- read.csv(here("data/poststratified/state_vote_for_biden.csv"))
@@
 For this analysis, we have used two datasets.
-The datasets were cleaned and analysed using the statistical programming software R [@citeR] along with the help tidyverse [@citeTidyverse], knitr [@citeKnitr], ggplot2 [@citeGgplot], here [@citeHere], ADD MORE
+The datasets were cleaned and analysed using the statistical programming software R [@citeR] along with the help tidyverse [@citeTidyverse], knitr [@citeKnitr], ggplot2 [@citeGgplot], here [@citeHere], dplyr [@citeDplyr], arrow [@citeArrow], rstanarm [@citeRstanarm], broom.mixed [@citeBroomMixed], modelsummary [@citeModelSummary] and kableExtra [@citeKableExtra].
 
@@
 ```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
 head(survey_data)
@@
 ```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
 head(poststrat_data)

### 2024-03-16 scripts/00-simulate_data.R

@@ -7,6 +7,8 @@
 # Pre-requisites: None
 
 #### Workspace setup ####
+# install.packages("tidyverse")
+# install.packages("dplyr")
 library(tidyverse)
 library(dplyr)
 
@@ -34,6 +36,26 @@ survey_data <- data.frame(age, sex, income, state, vote_biden)
 head(survey_data)
 
 #### Simulate poststratification dataset ####
+set.seed(323)
+n <- 10000
+
+age <- sample(18:98, n, replace = TRUE) # Simulate age data (ranging from 18 to 98)
+sex <- sample(c("Female", "Male"), n, replace = TRUE) # Simulate sex data (female or male)
+income <- round(runif(n, min = 20000, max = 1000000)) # Simulate family income data (ranging from 20000 to 1000000)
+
+state_labels <- c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "Rhode Island",
+                  "Vermont", "Delaware", "New Jersey", "New York", "Pennsylvania", "Illinois", "Indiana", "Michigan", "Ohio",
+                  "Wisconsin", "Iowa", "Kansas", "Minnesota", "Missouri", "Nebraska", "North Dakota", "South Dakota", "Virginia",
+                  "Alabama", "Arkansas", "Florida", "Georgia", "Louisiana", "Mississippi", "North Carolina", "South Carolina",
+                  "Texas", "Kentucky", "Maryland", "Oklahoma", "Tennessee", "West Virginia", "Arizona", "Colorado", "Idaho", "Montana",
+                  "Nevada", "New Mexico", "Utah", "Wyoming", "California", "Oregon", "Washington", "Alaska", "Hawaii", "Puerto Rico",
+                  "State groupings (1980 Urban/rural sample)", "Overseas Military Installations", "District of Columbia",
+                  "State not identified")
+state <- sample(state_labels, n, replace = TRUE) # Simulate state data
+
+
+poststratification_data <- data.frame(age, sex, income, state)
+head(poststratification_data)
 
 
 

### 2024-03-16 scripts/03-model.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Multilinear logistic regression model
 # Author: Kaavya Kalani, Monica Sainani
-# Date: 9 March 2024
+# Date: 11 March 2024
 # Contact: kaavya.kalani@mail.utoronto.ca, monica.sainani@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Run 01-data_cleaning.R
@@ -35,26 +35,6 @@ single_bay <-
 
 summary(single_bay)
 
-# multi_bay <-
-#   stan_glmer(
-#     vote_biden ~ sex + (1|age) + (1|state) + (1|education) + (1|income),
-#     data = survey_data,
-#     family = binomial(link = "logit"),
-#     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-#     prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-#     cores = 4,
-#     adapt_delta = 0.99,
-#     seed = 853
-#   )
-# 
-# summary(multi_bay)
-# 
-# freq <-  
-#   glm(vote_biden ~ as.factor(sex) + as.factor(age) + as.factor(state) + as.factor(education) + as.factor(income),
-#                data = survey_data,
-#                family = "binomial")
-#   
-# summary(freq)
 
 #### Save model ####
 saveRDS(
@@ -62,13 +42,4 @@ saveRDS(
   file = "models/single_bay.rds"
 )
 
-# saveRDS(
-#   multi_bay,
-#   file = "models/multi_bay.rds"
-# )
-
-# saveRDS(
-#   freq,
-#   file = "models/freq.rds"
-# )
 

### 2024-03-16 scripts/04-poststratification.R

@@ -1,29 +1,20 @@
 #### Preamble ####
 # Purpose: Poststratification for the model
 # Author: Kaavya Kalani, Monica Sainani
-# Date: 9 March 2024
+# Date: 13 March 2024
 # Contact: kaavya.kalani@mail.utoronto.ca, monica.sainani@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Run 04-model.R
 
 
 #### Workspace setup ####
-
-library(gutenbergr)
-library(haven)
-library(knitr)
-library(labelled)
-library(modelsummary)
-library(rstanarm)
-library(tidybayes)
-library(broom.mixed)
-library(here)
+# install.packages("tidyverse")
 library(tidyverse)
 
-# read the datasets and model
-poststrat_data <- arrow::read_parquet(file = here("data/analysis_data/poststrat_data.parquet"))
-survey_data <- arrow::read_parquet(file = here("data/analysis_data/survey_data.parquet"))
-us_election_model <- readRDS(file = here("models/single_bay.rds"))
+# Read the datasets and model
+poststrat_data <- arrow::read_parquet("data/analysis_data/poststrat_data.parquet")
+survey_data <- arrow::read_parquet("data/analysis_data/survey_data.parquet")
+us_election_model <- readRDS("models/single_bay.rds")
 
 #### Post-stratification 1 (State) ####
 
@@ -180,11 +171,11 @@ general_data <- survey_data %>%
   mutate(prop = n / sum(n))
 
 #### Save all tables ####
-write.csv(state_vote_for_biden, file = "data/poststratified/state_vote_for_biden.csv", row.names = FALSE)
-
-write.csv(age_vote_for_biden, file = "data/poststratified/age_vote_for_biden.csv", row.names = FALSE)
-
-write.csv(sex_vote_for_biden, file = "data/poststratified/sex_vote_for_biden.csv", row.names = FALSE)
-
-write.csv(general_data, file = "data/poststratified/general_vote_for_biden.csv", row.names = FALSE)
+# write.csv(state_vote_for_biden, file = "data/poststratified/state_vote_for_biden.csv", row.names = FALSE)
+# 
+# write.csv(age_vote_for_biden, file = "data/poststratified/age_vote_for_biden.csv", row.names = FALSE)
+# 
+# write.csv(sex_vote_for_biden, file = "data/poststratified/sex_vote_for_biden.csv", row.names = FALSE)
+# 
[…]

### 2024-03-16 paper/paper.qmd
Visualisation added

@@
 
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-cap: Vote count distribution in survey data
+#| label: vote-biden
+
+ggplot(survey_data, aes(x = as.factor(vote_biden))) +
+  geom_bar(fill = "skyblue", color = "black") +
+  labs(x = "Party Nominee",
+       y = "Count") +
+  scale_x_discrete(labels = c("Vote for Biden", "Vote for Trump"))
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-cap: Variables of interest and the vote proportion distribution
+#| fig-subcap: ["Age", "Sex", "Income", "Education", "State"]
+#| label: fig-variables
+#| layout-ncol: 2
+
+# Calculate the proportion of individuals who voted for Biden for each age
+age_proportions <- survey_data %>%
+  group_by(age, vote_biden) %>%
+  summarize(count = n()) %>%
+  mutate(proportion = count / sum(count))
+
+ggplot(age_proportions, aes(x = age, y = proportion, color = as.factor(vote_biden))) +
+  geom_point(position = position_dodge(width = 0.5), size = 3) +
+  geom_line(aes(group = vote_biden), position = position_dodge(width = 0.5)) +
+  labs(x = "Age Category", y = "Proportion", color = "Vote") +
+  scale_color_manual(values = c("blue", "red"), labels = c("Vote for Trump", "Vote for Biden")) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+# Calculate the proportion of individuals who voted for Biden for each sex
+sex_proportions <- survey_data %>%
+  group_by(sex, vote_biden) %>%
+  summarize(count = n()) %>%
+  mutate(proportion = count / sum(count))
+
+ggplot(sex_proportions, aes(x = sex, y = proportion, color = as.factor(vote_biden))) +
+  geom_point(position = position_dodge(width = 0.5), size = 3) +
+  geom_line(aes(group = vote_biden), position = position_dodge(width = 0.5)) +
+  labs(x = "Sex", y = "Proportion", color = "Vote") +
+  scale_color_manual(values = c("blue", "red"), labels = c("Vote for Trump", "Vote for Biden")) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+# Calculate the proportion of individuals who voted for Biden for each income level
+income_proportions <- survey_data %>%
+  group_by(income, vote_biden) %>%
+  summarize(count = n()) %>%
+  mutate(proportion = count / sum(count))
+
+ggplot(income_proportions, aes(x = income, y = proportion, color = as.factor(vote_biden))) +
+  geom_point(position = position_dodge(width = 0.5), size = 3) +
+  geom_line(aes(group = vote_biden), position = position_dodge(width = 0.5)) +
+  labs(x = "Income", y = "Proportion", color = "Vote") +
+  scale_color_manual(values = c("blue", "red"), labels = c("Vote for Trump", "Vote for Biden")) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+# Calculate the proportion of individuals who voted for Biden for each education level
+education_proportions <- survey_data %>%
+  group_by(education, vote_biden) %>%
+  summarize(count = n()) %>%
+  mutate(proportion = count / sum(count))
+
+ggplot(education_proportions, aes(x = education, y = proportion, color = as.factor(vote_biden))) +
+  geom_point(position = position_dodge(width = 0.5), size = 3) +
+  geom_line(aes(group = vote_biden), position = position_dodge(width = 0.5)) +
+  labs(x = "Education", y = "Proportion", color = "Vote") +
+  scale_color_manual(values = c("blue", "red"), labels = c("Vote for Trump", "Vote for Biden")) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+# Calculate the proportion of individuals who voted for Biden for each state
+state_proportions <- survey_data %>%
+  group_by(state, vote_biden) %>%
+  summarize(count = n()) %>%
+  mutate(proportion = count / sum(count))
+
+ggplot(state_proportions, aes(x = state, y = proportion, color = as.factor(vote_biden))) +
+  geom_point(position = position_dodge(width = 0.5), size = 3) +
+  geom_line(aes(group = vote_biden), position = position_dodge(width = 0.5)) +
+  labs(x = "State", y = "Proportion", color = "Vote") +
+  scale_color_manual(values = c("blue", "red"), labels = c("Vote for Trump", "Vote for Biden")) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+```
+
 ## Post-stratification Data
@@
 These variables were cleaned and the dataset was preprocessed so the values and categories match with the survey dataset.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-cap: Comparison of Survey and Census Data Proportions
+#| fig-subcap: ["Age", "Sex", "Income", "Education", "State"]
+#| label: fig-proportions
+#| layout-ncol: 2
+
+library(dplyr)
+library(ggplot2)
+library(reshape2)
+
+# Group and summarize survey_data by age
+df_survey <- survey_data %>% 
+  count(age) %>%
+  mutate(sur_prop = n / sum(n)) %>%
+  select(age, sur_prop)
+
+# Group and summarize poststrat_data (census data) by age
+df_census <- poststrat_data %>% 
+  count(age) %>% 
+  mutate(census_prop = n / sum(n)) %>%
+  select(age, census_prop)
+
+# Merge the two datasets
+combined_data <- left_join(df_census, df_survey, by = "age")
+
+# Plotting
+ggplot(combined_data, aes(age)) +
+  geom_point(aes(y = census_prop, color = "Census data"), size = 3) +
+  geom_point(aes(y = sur_prop, color = "Survey data"), size = 3) +
+  scale_color_manual(values = c("Census data" = "salmon", "Survey data" = "skyblue")) +
+  labs(title = "Age",
+       x = "Age",
+       y = "Proportion") +
+  theme_minimal()
+
+# Group and summarize survey_data by sex
+df_survey <- survey_data %>% 
+  count(sex) %>%
+  mutate(sur_prop = n / sum(n)) %>%
+  select(sex, sur_prop)
+
+# Group and summarize poststrat_data (census data) by sex
+df_census <- poststrat_data %>% 
+  count(sex) %>% 
+  mutate(census_prop = n / sum(n)) %>%
+  select(sex, census_prop)
+
+# Merge the two datasets
+combined_data <- left_join(df_census, df_survey, by = "sex")
+
+# Plotting
+ggplot(combined_data, aes(sex)) +
+  geom_point(aes(y = census_prop, color = "Census data"), size = 3) +
+  geom_point(aes(y = sur_prop, color = "Survey data"), size = 3) +
+  scale_color_manual(values = c("Census data" = "salmon", "Survey data" = "skyblue")) +
+  labs(title = "Sexs",
+       x = "Sex",
+       y = "Proportion") +
+  theme_minimal()
+
+# Group and summarize survey_data by income
+df_survey <- survey_data %>% 
+  count(income) %>%
+  mutate(sur_prop = n / sum(n)) %>%
+  select(income, sur_prop)
+
+# Group and summarize poststrat_data (census data) by income
+df_census <- poststrat_data %>% 
+  count(income) %>% 
+  mutate(census_prop = n / sum(n)) %>%
+  select(income, census_prop)
+
+# Merge the two datasets
+combined_data <- left_join(df_census, df_survey, by = "income")
+
+# Plotting
+ggplot(combined_data, aes(income)) +
+  geom_point(aes(y = census_prop, color = "Census data"), size = 3) +
+  geom_point(aes(y = sur_prop, color = "Survey data"), size = 3) +
+  scale_color_manual(values = c("Census data" = "salmon", "Survey data" = "skyblue")) +
+  labs(title = "Income",
+       x = "Income",
+       y = "Proportion") +
+  theme_minimal() + 
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+# Group and summarize survey_data by education
+df_survey <- survey_data %>% 
+  count(education) %>%
+  mutate(sur_prop = n / sum(n)) %>%
+  select(education, sur_prop)
+
+# Group and summarize poststrat_data (census data) by education
+df_census <- poststrat_data %>% 
+  count(education) %>% 
+  mutate(census_prop = n / sum(n)) %>%
+  select(education, census_prop)
+
+# Merge the two datasets
+combined_data <- left_join(df_census, df_survey, by = "education")
+
+# Plotting
+ggplot(combined_data, aes(education)) +
+  geom_point(aes(y = census_prop, color = "Census data"), size = 3) +
+  geom_point(aes(y = sur_prop, color = "Survey data"), size = 3) +
+  scale_color_manual(values = c("Census data" = "salmon", "Survey data" = "skyblue")) +
+  labs(title = "Education",
+       x = "Education",
+       y = "Proportion") +
+  theme_minimal() + 
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+# Group and summarize survey_data by state
+df_survey <- survey_data %>% 
+  count(state) %>%
+  mutate(sur_prop = n / sum(n)) %>%
+  select(state, sur_prop)
+
+# Group and summarize poststrat_data (census data) by state
+df_census <- poststrat_data %>% 
+  count(state) %>% 
+  mutate(census_prop = n / sum(n)) %>%
+  select(state, census_prop)
+
+# Merge the two datasets
+combined_data <- left_join(df_census, df_survey, by = "state")
+
+# Plotting
+ggplot(combined_data, aes(state)) +
+  geom_point(aes(y = census_prop, color = "Census data"), size = 3) +
+  geom_point(aes(y = sur_prop, color = "Survey data"), size = 3) +
+  scale_color_manual(values = c("Census data" = "salmon", "Survey data" = "skyblue")) +
+  labs(title = "State",
+       x = "State",
+       y = "Proportion") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+```
 
@@
 ```
-
 @tbl-model-coefficients shows the coefficients for our model that can be used in the equation along with the standard error and the 95% confidence interval.
@@
 ```
-@fig-support-state
 In @fig-support-state we see the percentage of the votes for Biden/Democratic party. This graph uses the data we

### 2024-03-16 paper/paper.qmd
Slight change in dataset and model

@@
 We find a strong relationship between age, family income, education, state of residency, sex and likely voting outcome.
-Our central forecast is a Biden re-election with a 10% lead.
+Our central forecast is a Biden re-election.
 This research provides insights into the factors affecting political engagement in the U.S., potentially guiding future electoral strategies and policy discussions."
@@
 For the variable we are predicting, we created a binary variable which is 1 if the panelist favors Biden and 0 if the panelist favors Trump.
-There was no question in the poll which asked this direct question so we used “Overall, do you have a favorable or unfavorable impression of...Joe Biden?”.
-We acknowledge that not favoring Joe Biden does not necessarily imply favoring Donald Trump, but since they are the two major potential candidates, we classify the panelists who chose “Unfavorable” as supporting Trump, i.e not supporting Biden.
+There was no question in the poll which asked this direct question so we used “Overall, do you have a favorable or unfavorable impression of...Donald Trump?”.
+We acknowledge that not favoring Donald Trump does not necessarily imply favoring Joe Biden, but since they are the two major potential candidates, we classify the panelists who chose “Unfavorable” as supporting Biden, i.e not supporting Trump.
 
@@
        y = "Count") + 
-  scale_x_discrete(labels = c("Vote for Biden", "Vote for Trump"))
+  scale_x_discrete(labels = c("Vote for Trump", "Vote for Biden"))
 ```
+
+@vote-biden shows the distribution of vote count distribution our survey data. We see that
 
@@
 
+$$
+\begin{aligned}
+\beta_0 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_1 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_2 & \sim \mbox{Normal}(0, 2.5)\\
+\beta_3 & \sim \mbox{Normal}(0, 2.5)
+\end{aligned}
+$$
+
 where,
@@
 
-In our model, normal priors with a mean of 0 and a standard deviation of 2.5 are used for both the coeffients and the intercept.
+In our model, normal priors with a mean of 0 and a standard deviation of 2.5 are used for both the coefficients and the intercept.
 Setting the mean of the priors to 0 implies that there is no expectation of a particular direction or magnitude for the coefficients or intercept.
@@
        x = "Vote") +
-  scale_x_discrete(labels = c("Vote for Biden", "Will not vote for Biden"))
+  scale_x_discrete(labels = c("Vote for Trump", "Vote for Biden"))
 

### 2024-03-16 scripts/01-data_cleaning.R

@@ -20,7 +20,7 @@ poststrat_data <- arrow::read_parquet("data/raw_data/usa.parquet")
 
 # Clean and preprocess survey data
 survey_data <- survey_data %>% 
-  filter(Q1_2 %in% c("Favorable", "Unfavorable")) %>% 
+  filter(Q1_1 %in% c("Favorable", "Unfavorable")) %>% 
   mutate(age = case_when(
             ppage <= 29 ~ "18-29",
             ppage <= 44 ~ "30-44",
@@ -39,8 +39,8 @@ survey_data <- survey_data %>%
            ppgender == "Male" ~ "male",
            TRUE ~ as.character(ppgender)),
          vote_biden = case_when(
-           Q1_2 == "Favorable" ~ 1,
-           Q1_2 == "Unfavorable" ~ 0)) %>%
+           Q1_1 == "Favorable" ~ 0,
+           Q1_1 == "Unfavorable" ~ 1)) %>%
   select(age, education, income, state, sex, vote_biden) 
 
 # Clean and preprocess poststratification data

### 2024-03-16 scripts/04-poststratification.R

@@ -171,11 +171,11 @@ general_data <- survey_data %>%
   mutate(prop = n / sum(n))
 
 #### Save all tables ####
-# write.csv(state_vote_for_biden, file = "data/poststratified/state_vote_for_biden.csv", row.names = FALSE)
-# 
-# write.csv(age_vote_for_biden, file = "data/poststratified/age_vote_for_biden.csv", row.names = FALSE)
-# 
-# write.csv(sex_vote_for_biden, file = "data/poststratified/sex_vote_for_biden.csv", row.names = FALSE)
-# 
-# write.csv(general_data, file = "data/poststratified/general_vote_for_biden.csv", row.names = FALSE)
+write.csv(state_vote_for_biden, file = "data/poststratified/state_vote_for_biden.csv", row.names = FALSE)
+
+write.csv(age_vote_for_biden, file = "data/poststratified/age_vote_for_biden.csv", row.names = FALSE)
+
+write.csv(sex_vote_for_biden, file = "data/poststratified/sex_vote_for_biden.csv", row.names = FALSE)
+
+write.csv(general_data, file = "data/poststratified/general_vote_for_biden.csv", row.names = FALSE)
 

### 2024-03-16 paper/paper.qmd
Descriptions added

@@
 
-@vote-biden shows the distribution of vote count distribution our survey data. We see that
+@vote-biden shows the distribution of vote count in our survey data.
+We see that the survey panelists are heavily in favor of Joe Biden with a over twice of the panelists not favoring Trump.
+The bias can be a potential weakness of the dataset and we aim to uncover the factors influencing these political preferences.
 
@@
 ```
+@fig-variables shows the vote proportion distribution for our chosen variables of interests.
+We notice a noticeable difference in the proportions that favor Trump or Biden.
+Some striking trends we notice in our survey data are more male panelists supporting Trump, majority of Trump supporters being in the 45-59 age group and Trump support declining with increase in education levels.
+While most states seem to support Trump, we see that some states support Trump over Biden.
+This is a fresh change to see over all other factors showing a clear preference for Biden in all categories.
 
@@
   scale_color_manual(values = c("Census data" = "salmon", "Survey data" = "skyblue")) +
-  labs(title = "Sexs",
+  labs(title = "Sex",
        x = "Sex",
@@
 ```
+@fig-proportions show a comparison of the survey and census data proportions.
+We notice that for all the variables of interest, the survey is not entirely representative of the population.
+For example, we see that females are underrepresented in the survey with the survey having about 49% females whereas the census shows 51% females.
+We also see how population with highest education attained being high school are under-represented whereas population with highest education attained being more than high school are over-representated.
+Difference in proportions like these can be seen in all the variables of interests and this is why post-stratifying is important to ensure precision and better representation of the population.
 
@@
 In @fig-support-state we see the percentage of the votes for Biden/Democratic party. This graph uses the data we
-collected after the post-stratification. Here, the red dots on the graph represent the voting percentage and the red
-lines are the error bars. The black vertical line with dashes indicates the line of proportion. Here, it is the equal
-split between support for Biden and support for the opposing candidate, Trump.
+collected after the post-stratification.
+Here, the red dots on the graph represent the voting percentage and the red lines are the error bars.
+The black vertical line with dashes indicates the line of proportion.
+Here, it is the equal split between support for Biden and support for the opposing candidate, Trump.
 
@@
 are found from the Northeast,Midwest  and the West Coast states. For example, the difference in votes for Biden
-between Oklahoma and New Mexico is more than 50%, but states like Oregon and Maine are neutral in terms of voting
-for either parties.
+between Oklahoma and New Mexico is more than 50%, but states like Oregon and Maine are neutral in terms of voting for either parties.
 
@@
 
-In @fig-support-age, we used the data to find the proportion of the voters categorized by age, who would vote for the
-Democratic party. We found a pattern that the group of youngest voters and the group of the senior citizen votes are
-more likely to vote for the Democratic Party(Biden).
-
-We see that 45% of the individuals from age group 18-29 are voting
-for biden as well as 50% of the individuals from the age group of 60+ years. Meanwhile, only 35% of people in the age
-group of 45-59 are voting for Biden. Hence, in this graph we see the trend that people of ages from 30-59 are less likely
-to vote for the democratic party.
+In @fig-support-age, we used the data to find the proportion of the voters categorized by age, who would vote for the Democratic party.
+We see that the younger generation are more likely to vote for Biden compared to the older.
+Americans under the age of 45 are 70% likely to vote for Trump.
+The support for Trump drops to around 60% in the 45-59 age group.
+While this is favorable towards Biden, it does indicate that more people between 45 to 59 years are likely to vote for Trump compared to other age groups.
 
@@
 ```
-
-@fig-support-sex shows the proportion of Female voters that would vote for the Democratic Party as compared to the Male
-voters. We see more than 10% difference in the proportion. Here, we see that Women are 10% more likely to vote for Biden.
-While only 40% of the men in US would vote for Biden, more than 50% of women would vote for the Biden.
+@fig-support-sex shows the proportion of Female voters that would vote for the Democratic Party as compared to the Male voters.
+We see a 10% difference in the proportion where women are 10% more likely to vote for Biden.
 
@@
 
-@fig-support-general looks at the division of voters for each party. The sample here is all the individuals who took
-the survey and what the proportions are in general population of the United States. We see that more than 50% of the
-population would vote for Biden and the rest would vote for Trump. This graph clearly shows the
-conclusion we can deduce from our research that with this survey done in 2022, it is clear that the population of the
-United States is inclined towards the Democratic Party.
+@fig-support-general looks at the division of voters for each party as an overall.
+We see that even after applying post-stratification, the preference has only changed by a negligible amount with a large majority of the population likely to vote for Trump.
+This graph clearly shows the conclusion we can deduce from our research that it is clear that the population of the United States is more inclined towards the Democratic Party.
 

### 2024-03-16 paper/paper.qmd
Reference issue fixed

@@
 #| fig-cap: Vote count distribution in survey data
-#| label: vote-biden
+#| label: fig-vote-biden
 
@@
 
-@vote-biden shows the distribution of vote count in our survey data.
+@fig-vote-biden shows the distribution of vote count in our survey data.
 We see that the survey panelists are heavily in favor of Joe Biden with a over twice of the panelists not favoring Trump.

### 2024-03-16 paper/paper.qmd
other
Cleaning instructions added in appendix

@@
 # Introduction
-A key turning point in the political environment of the United States occurred with the 2020 presidential election,
-which was marked by extreme conflict and serious economic and social challenges. We analyze data from the American
-Community Survey (ACS) 2022 sample, which offers a thorough understanding of demographic, socioeconomic,
-and geographic aspects across the nation, in order to comprehend the patterns underlying voting choices during
-this 2024 election period. This paper focuses on the two major political parties, Democratic and Republican.
-
-The rivalry between the Democratic and Republican parties, as embodied by their opposing candidates,
-Joe Biden and Donald Trump, has been an iconic feature of the modern American political scene. The Democratic Party's
-nominee is the veteran politician and former Obama vice president Joe Biden. With an emphasis on middle-class families,
-Biden's program prioritizes measures that advance racial justice, healthcare reform, climate action, and economic
-regeneration. He is an advocate of a more compassionate and inclusive style of governing, hoping to bring the
-country back together and bring decorum back into political conversation.
-
-On the other hand, the current Republican Party standard-bearer and President Donald Trump represents a more populist
-and nationalist agenda. A major focus on economic nationalism, immigration restrictions, tax cuts, and
-deregulation have been highlights of Trump's leadership. His administration prioritized themes like law and order,
-trade protectionism, and border security as part of its conservative agenda. Trump's aggressive approach to politics
-is a defining characteristic of his leadership style; it frequently divides public opinion and provokes strong emotions
-from both supporters and opponents.
-
-Our estimand is the relationship between demographic, socio-economic, and geographic variables (such as age, income,
-education, state of residence, and gender) and voting behavior during the 2022 presidential election.
-Using this extensive dataset, our goal is to identify trends and factors that influence voting behaviour,
-clarifying the complex relationship between personal traits and local factors that influences election results.
-
-Our research adds to a greater understanding of the factors influencing voting decisions through comprehensive
-statistical analysis, yielding insightful data that may influence future election campaigns and policy discussions.
-Through highlighting the subtleties of voting patterns in the 2022 election, our research provides a critical insight
-on the condition of American democracy and the factors motivating involvement in politics and engagement.
-
-The paper is further organised into three sections: Data, Model, Results, and Discussion. In the Data section, we
-discuss how the datasets to be used for the analysis were obtained and pre-processed. We will explain the variables of
-interest in the datasets for the analysis. The Model section describes the model being used for the analysis. The Results
-section will then highlight and discuss the trends and associations found during the analysis. Lastly, the Discussion
-section will talk about some interesting trends found in Results in depth, link it to the real world and also highlight
-the weaknesses of our analysis.
 
@@
 ```
+
 In @fig-support-state we see the percentage of the votes for Biden/Democratic party. This graph uses the data we
@@
 
-# Discussion
-
-## Insights Gained:
-
-Our analysis, based on sample data from the 2022 survey, followed by post-stratification focusing on key variables, Age,
-State, and Sex, yielded valuable insights into voter behavior. Notably, our findings indicated an inclination within the
-population towards voting for Biden from the Democratic Party.
-
-In the graph showcasing different regions and their voting for Biden, we noticed that
-this is similar to the findings in an article by statistician James E. Campbell,  [@James_2016] that address the same prediction for 2024. Our model and the
-article predict that states that support Trump are from the middle and Midwest parts of the US, whereas coastal states
-and the Northeast tend to vote for Biden.
-
-Historically, rural, conservative-leaning towns with strong cultural ties to traditional values and industries like
-manufacturing and agriculture have defined the middle and Midwest regions of the United States.
-These areas frequently support gun rights, minimal government intrusion, and religious liberty—issues that are strongly
-aligned with the agenda of the Republican Party, which Trump advocated. In contrast, the Northeast and coastline states,
-which are renowned for their progressive views on culture, urban facilities, and diverse populations, have historically
-tended toward more liberal policies and social ideals that are more in line with the Democratic Party platform, which Biden advocated.
-
-We also created a graph highlighting different age groups and what their voting opinions are. We noticed that
-the youngest group of voters (age 18-24). The emphasis on social justice and LGBTQ+ rights among the younger generation
-may be the cause of this difference[@Brownstein_2023]. Since there are undoubtedly more young people who identify as sexual minorities,
-there is a trend for them to support Biden and the Democratic Party. We also noticed that the older generation i.e
-individuals of age group higher than 60 years old were inclined to voting for Biden as well.
-National security and conservative principles, which are frequently connected to the Republican Party,
-may have an impact on the elder generation.
-
-Lastly, our third variable of study was Sex. We noticed a slight bias towards the democratic party from the females than
-the male gender. Women are predicted to vote for the democratic party 0.6 times more than men.
-Studies have consistently demonstrated that there is a gender difference in political ideology, with women generally
-favoring liberal or progressive policies over those of males.[@Igielnik_2021]
-This discrepancy is sometimes linked to the fact that women prioritize issues more than men do, with women giving social
-welfare, healthcare, education, and reproductive rights—all of which are more in line with the Democratic Party program.
-With its focus on women's rights, healthcare reform, and inclusivity, Biden's campaign may have connected more
-deeply with female voters, resulting in higher levels of support.
-
-## Comparative Analysis:
-Our findings align with existing literature on electoral behavior, which underscores the significance of demographic
-variables in shaping voter preferences. While our study offers insights into the inclination towards voting for Biden,
-it is important to contextualize our findings within the broader landscape of political science research.[Gelman]
-
-Our analysis contributes to a growing body of evidence highlighting the complex interplay between socio-economic factors
-and political attitudes, providing valuable insights for future research and decision-making in the political arena.
-
-## Contributions to Knowledge:
-By focusing our analysis on specific demographic variables, we contributed to a deeper understanding of the underlying
-determinants shaping electoral outcomes. Our study corroborates existing literature suggesting that demographic factors
-such as age, income, education, and geographic location play pivotal roles in influencing voter preferences. [@Gelman]
-Moreover, our utilization of post-stratification techniques allowed for a more granular examination of voting behavior,
-offering insights into the heterogeneity of voter sentiment across different demographic groups.
-
-## Weaknesses and Limitations:
-
-While our analysis provides valuable insights into voter behavior, it is important to acknowledge several limitations. 
-Firstly, our reliance on sample data from the 2022 survey may introduce biases and inaccuracies in our predictions.
-Our reliance on sample data from the 2022 survey may introduce a sampling bias, as the sample might not fully represent
-the diversity of the electorate. Consequently, this could lead to inaccuracies in our predictions and limit the
-generalizability of our findings to the broader population.
-
-Additionally, while we focused on key demographic variables, other influential factors such as political ideology and
-campaign messaging were not fully accounted for in our model. This means that our research has the inaccuracy of
-Incomplete Variable Consideration[@Funakoshi_2020].
-
-Furthermore, uncertainties in the post-stratification process may impact the reliability of our estimates,
-warranting caution in interpreting the observed inclination towards voting for Biden.
-
-## Future Directions:
-Moving forward, further research is warranted to address the limitations identified in our study and advance our
-understanding of electoral dynamics. Future studies could explore the integration of additional data sources,
-such as social media sentiment analysis and polling data, to enhance the accuracy of election predictions.
-Additionally, efforts to improve transparency and reproducibility in data collection and analysis are essential
-for enhancing the reliability and validity of electoral forecasts.
-
 # Appendix
@@
 ## Cleaning
+For the survey dataset, the cleaning steps we took were:
+
+1.
+Initial cleaning involved filtering the dataset to retain only responses relevant to the analysis.
+Responses were filtered based on the participant's stance on Q1_1 which is “Overall, do you have a favorable or unfavorable impression of...Donald Trump?”, specifically selecting those labeled as "Favorable" or "Unfavorable".
+
+2. Age groups were defined based on the participants' reported age (ppage). Participants were categorized into four age groups: "18-29", "30-44", "45-59", and "60+". The final was named age.
+
+3. Respondents' income information (ppinc7) was retained in its original format to provide insights into income distribution. The final column was named income.
+
+4.
+Education levels were standardized to facilitate analysis.
+Respondents' education information (ppeducat) was categorized into three groups: "High school", "More than high school" (including "Bachelors degree or higher" and "Some college"), and "Less than high school".
+This was done to match the education level separation available in the poststratification dataset.
+The final column was named education.
+
+5. The state of residence for each participant (ppstaten) was preserved for regional analysis. The final column was named state.
+
+6.
+Gender information (ppgender) was just had a change of turning the value to lower case.
+The final column was named sex as just having these two criterias implies that and to maintain with the poststratification data reporting sex as well.
+
+7.
+We created a binary variable for favorability of Biden.
+Responses from Q1_1 labeled as "Favorable" were coded as 0, while responses labeled as "Unfavorable" were coded as 1.
+The final column was named vote_biden.
+
+8. The final survey dataset contained the above mentioned 6 columns: age, education, income, state, sex, and the coded variable for Biden vote.
+
+For the poststratification dataset, the cleaning steps we took were:
+
+1.
+Initially, the "AGE" variable was converted to numeric format.
+Subsequently, data points were filtered to retain only those individuals aged 18 and above, excluding individuals from Vermont and Montana and cases where total household income (FTOTINC) was recorded as less than zero.
+Individuals from Vermont and Montana were excluded from the dataset to maintain consistency with the survey data as the survey data does not involve any records from these two states.
+
+2. Participants' ages (AGE) were categorized into four groups: "18-29", "30-44", "45-59", and "60+". The final column was named age.
+
+3.
+Education levels were standardized into four categories: "High school", "More than high school", and "Less than high school".
+This categorization was based on the reported education level (EDUC), with corresponding numerical codes for each category.
+The final column was named education.
+
+4. States were labeled according to their corresponding numerical codes (STATEICP), assigning each code to its respective state name. The final column was named education.
+
+5. Sex information (SEX) was changed for the codes to their respective values: "male" and "female". The final column was named sex.
+
+6.
+Household income (FTOTINC) was categorized into the same division of income brackets as the survey data: Less than $10,000, $10,000 to $24,999, $25,000 to $49,999, $50,000 to $74,999, $75,000 to $99,999, $100,000 to $149,999 and $150,000 or more.
+The final column was named income.
+
+7. The final poststratification dataset contained the above mentioned 5 columns: age, education, income, state and sex.
+
+Now we are ensured that both the datasets have the same variable names and respective categories.
 

### 2024-03-16 paper/paper.qmd
push the graph change- for states

@@
   labs(
-    x = "State",
-    y = "Estimated proportion support for Biden",
+    x = "Estimated proportion support for Biden",
+    y = "State",
     color = "Source"
@@
   coord_flip() +
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom") +
+  theme(axis.text.y = element_text(size = 7, margin = margin(0, 10, 0, 10))) # Adjust size of y-axis text
 

### 2024-03-16 paper/paper.qmd
state graph

@@
 ```
-
 In @fig-support-state we see the percentage of the votes for Biden/Democratic party. This graph uses the data we
@@
 The black vertical line with dashes indicates the line of proportion.
-Here, it is the equal split between support for Biden and support for the opposing candidate, Trump.
+Using that, we see that majority of the states of United States are in support for Biden and only 4 states are in support for the opposing candidate, Trump.
 
@@
 and the states on the right are voting for Biden. We see that the majority of the votes for the democratic party
-are found from the Northeast,Midwest  and the West Coast states. For example, the difference in votes for Biden
-between Oklahoma and New Mexico is more than 50%, but states like Oregon and Maine are neutral in terms of voting for either parties.
+are found from the Midwest  and the Southern states while only Rhode Island, Oklahoma, Missouri and Nevada are inclined to vote for Trump.
+Rhode Island is almost 75% more likely to vote for Trump, but in competition most of of Alaska, Hawaii, North Dakota, New Mexico and Kansas are inclined to vote for Biden.
 

### 2024-03-16 paper/paper.qmd
intro added

@@
 # Introduction
+
+The United States stands as one of the most influential political entities globally.
+This status implies that the leadership of the nation can cause indirect influence on global affairs [@Barnes_2021].
+Since its inaugural election in 1788, the US Presidential elections have been a quadrennial tradition, building the course of the nation's history every four years.
+This year, the 60th presidential election will be held on Tuesday, November 5, 2024.
+As the nation of the United States approaches their 2024 presidential election, the political landscape is likely to undergo significant shifts and intensification.
+
+The 2020 election brought upon extreme conflict, serious economic and social challenges which changed the societal dynamics for Americans [@Barnes_2021].
+Although the decisions are made every 4 years, the United States conducts surveys every year to determine the opinions of citizens and predict the outcome of the elections.
+In our paper, we conduct a statistical analysis to forecast the major decision by focusing on the two major political parties, Democratic and Republican.
+
+The rivalry between the Democratic and Republican parties, as embodied by their opposing candidates, Joe Biden and Donald Trump, has been an iconic feature of the modern American political scene.
+The Democratic Party's nominee is the veteran politician and current president Joe Biden.
+With an emphasis on middle-class families, Biden's program prioritizes measures that advance racial justice, healthcare reform, climate action, and economic regeneration.
+He is an advocate of a more compassionate and inclusive style of governing, hoping to bring the country back together and bring decorum back into political conversation.
+
+On the other hand, the current Republican Party standard-bearer and former president
+Donald Trump represents a more populist and nationalist agenda.
+A major focus on economic nationalism, immigration restrictions, tax cuts, and deregulation have been highlights of Trump's leadership.
+His administration prioritized themes like law and order, trade protectionism, and border security as part of its conservative agenda.
+Trump's aggressive approach to politics is a defining characteristic of his leadership style; it frequently divides public opinion and provokes strong emotions from both supporters and opponents.
+
+Our estimand is the relationship between demographic, socio-economic, and geographic variables (such as age, income, education, state of residence, and sex) and political preferences.
+Using this extensive dataset, our goal is to identify trends and factors that influence voting behavior, clarifying the complex relationship between personal traits and local factors that influences election results.
+Eventually, we aim to predict will Joe Biden win the US 2024 Presidential Elections or will it be Donald Trump.
+
+We use the survey data from the Ipsos Poll [Roper #31120637] along with census data from the American Community Survey (ACS) 2022 sample to understand voting preferences and predict the US 2024 Presidential Elections winner.
+This is done by leveraging a Bayesian Logistic Regression model with our survey data, which is then post-stratified with our census data.
+We evaluate voting preferences over some demographic, socioeconomic, and geographic aspects across the nation and then make a final prediction.
+
+Our analysis led to us finding that overall, the younger generations, females and people of the midwest and southern states are highly supportive of the Democratic Party led by Biden.
+Our overall forecast is a Joe Biden re-election in the 2024 US Presidential Elections.
+
+Our research adds to a greater understanding of the factors influencing voting decisions through comprehensive statistical analysis, yielding insightful data that may influence future election campaigns and policy discussions.
+Through highlighting the subtleties of voting patterns, our research provides a critical insight on the condition of American democracy and the factors motivating involvement in politics and engagement.
+
+The paper is further organized into four sections: Data, Model, Results, and Discussion.
+In the Data section, we discuss how the datasets to be used for the analysis were obtained and pre-processed.
+We will explain the variables of interest in the datasets for the analysis.
+The Model section describes the model being used for the analysis.
+The Results section will then highlight and discuss the trends and associations found during the analysis.
+Lastly, the Discussion section will talk about some interesting trends found in Results in depth, link it to the real world and also highlight the weaknesses and future of our analysis.
 

### 2024-03-16 paper/paper.qmd
discussion added

@@
 
+# Discussion
+Our analysis modeled the variables of interest from the survey data, followed by post-stratification focusing on key variables.
+We plotted the trends for the post-stratified voting preferences by age, state, and sex.
+These visualizations yielded valuable insights into voter behavior.
+Notably, our findings indicated an inclination within the population towards voting for Biden from the Democratic Party.
+
+## Votes By States
+In the graph, @fig-support-state, we notice that all states in the U.S.
+other than Rhode Island, Missouri, Oklahoma and Nevada are more inclined to vote for the Democratic Party.
+We see that as we go towards the Midwest and Southern Regions, the support for Biden increases as more than 50% of the state is likely to vote for them.
+This is similar to the findings in [@James_2016] that address the same prediction for 2024.
+Our model and the article predict that states that support Trump are from the middle and Midwest parts of the US, whereas coastal states and the Northeast tend to vote for Biden.
+
+Historically, rural, conservative-leaning towns with strong cultural ties to traditional values and industries like manufacturing and agriculture have defined the middle and Midwest regions of the United States.
+These areas frequently support gun rights, minimal government intrusion, and religious liberty—issues that are strongly aligned with the agenda of the Republican Party, which Trump advocated.
+In contrast, the Northeast and coastline states, which are renowned for their progressive views on culture, urban facilities, and diverse populations, have historically
+tended toward more liberal policies and social ideals that are more in line with the Democratic Party platform, which Biden advocated.
+
+## Age and Sex
+@fig-support-age showcases different age groups and what their voting preferences are.
+We noticed that the youngest group of voters (age 18-29) and mid-age adults (30-44) are more than 70% inclined to vote for Biden.
+The emphasis on social justice and LGBTQ+ rights among the younger generation may be the cause of this difference.
+The democratic party led by Biden prioritizes these issues and speaks for the justice of the minorities [@Brownstein_2023].
+Since there are more young people who identify as sexual minorities and they make up 10% of the U.S.
+population [@Meyer_2024], there is a trend for them to support Biden and the Democratic Party.
+
+The middle-aged generation i.e individuals of age group from 30-44 years old were inclined to voting for Biden as well.
+Middle-aged voters may align more closely with Biden's policy proposals on issues such as healthcare, economic recovery, climate change, and social justice [@Igielnik_2021] .
+
+Additionally, we notice almost a 10% decline in the votes for Biden from older age groups (45+ years old).
+The National security and conservative principles, which are frequently connected to the Republican Party, may have an impact on the elder generation.
+
+@fig-support-sex shows a slight bias towards the democratic party from the females than the male gender.
+Predictions suggest that women are 10% more likely to vote for the Democratic Party than men.
+Studies have consistently demonstrated that there is a gender difference in political ideology, with women generally favoring progressive policies over those of males [@Igielnik_2021].
+This discrepancy is sometimes linked to the fact that women prioritize some  issues more than men do, with women giving social welfare, healthcare, education, and reproductive rights—all of which are more in line with the Democratic Party program [@Igielnik_2021].
+With its focus on women's rights, healthcare reform, and inclusivity, Biden's campaign seems to have connected more deeply with female voters, resulting in higher levels of support.
+
+## General conclusion
+@fig-support-general shows that in general the population of the United States is a lot more inclined to vote for Biden.
+We see a drastic difference in opinion of almost 40% i.e.
+people are 40% more likely to vote for Biden than vote for Trump.
+This conclusion highlighted the final prediction from our analysis.
+Furthermore, it sheds light on the prevailing social values and the direction individuals envision for their society.
+Some of the key issues that Americans wanted to see change were the healthcare reforms, climate change actions and racial justice and social equity.
+They wanted better national response to the COVID-19 pandemic, more ambitious climate policies and investments in renewable energy and lastly, a demand for reforms addressing systemic racism and criminal justice [@Brownstein_2023].
+
+## Contributions to Knowledge:
+
+By focusing our analysis on specific demographic variables, we contributed to a deeper understanding of the underlying determinants shaping electoral outcomes.
+Our study corroborates existing literature suggesting that demographic factors such as age, income, education, and geographic location play pivotal roles in influencing voter preferences [@Gelman].
+
+The utilization of post-stratification techniques in our analysis allowed for a more granular examination of voting behavior, offering insights into the heterogeneity of voter sentiment across different demographic groups.
+
+Lastly, it highlights the high support of the Democratic Party led by Biden from the younger generations, females and people of the midwest and southern states.
+
 # Appendix

### 2024-03-16 paper/paper.qmd
PDF rendered

@@
   theme(legend.position = "bottom") +
-  theme(axis.text.y = element_text(size = 7, margin = margin(0, 10, 0, 10))) # Adjust size of y-axis text
+  theme(axis.text.y = element_text(size = 4, margin = margin(0, 10, 0, 10))) # Adjust size of y-axis text
 
@@
 
-## Votes By States
+## Political preference trends in different states 
 In the graph, @fig-support-state, we notice that all states in the U.S.
@@
 
-## Age and Sex
+## Political preference trends in different age groups and sex categories
 @fig-support-age showcases different age groups and what their voting preferences are.

### 2024-03-16 paper/paper.qmd
Discussion complete and typos fixed

@@
 title: "Biden Projected to Win the 2024 US Presidential Elections"
-subtitle: "Analysis using data from American Community Survey 2022 and Ipsos Poll 2023 levereging Bayesian Logistic Regression with Post-Stratification"
+subtitle: "Analysis using data from American Community Survey 2022 and Ipsos Poll 2023 leveraging Bayesian Logistic Regression with Post-Stratification"
 author:
@@
 abstract: "We analyze American voting behavior, focusing on demographic, socioeconomic, and geographic factors.
-We use a Bayesian Logistic Regression model with Post-Stratification to predict the election outcome between Biden and Trump.
+A Bayesian Logistic Regression model with Post-Stratification is used to predict the election outcome between Biden and Trump.
 We find a strong relationship between age, family income, education, state of residency, sex and likely voting outcome.
@@
 
-Our analysis led to us finding that overall, the younger generations, females and people of the midwest and southern states are highly supportive of the Democratic Party led by Biden.
+Our analysis led to us finding that overall, the younger generations, females, and people of the Midwest and southern states are highly supportive of the Democratic Party led by Biden.
 Our overall forecast is a Joe Biden re-election in the 2024 US Presidential Elections.
@@
 The survey is conducted by Ipsos and the sponsor is ABC News.
-This poll is based on a nationally representative probability sample of 949 adults age 18 or older with oversamples among Black and Hispanic respondents.
+This poll is based on a nationally representative probability sample of 949 adults aged 18 or older with oversamples among Black and Hispanic respondents.
 The poll is a collection of demographic information and political preferences.
@@
 They were given the option between: Less than high school, High school, Some college and Bachelor's degree or higher.
-We then classified these into three categories:  Less than high school, High school,  More than high school
+We then classified these into three categories:  Less than high school, High school,  More than high school.
 
@@
 
-- Sex is the sex reported by the panelist and it is either male or female
+- Sex is the sex reported by the panelist and it is either male or female.
 
@@
 There was no question in the poll which asked this direct question so we used “Overall, do you have a favorable or unfavorable impression of...Donald Trump?”.
-We acknowledge that not favoring Donald Trump does not necessarily imply favoring Joe Biden, but since they are the two major potential candidates, we classify the panelists who chose “Unfavorable” as supporting Biden, i.e not supporting Trump.
+We acknowledge that not favoring Donald Trump does not necessarily imply favoring Joe Biden, but since they are the two major potential candidates, we classify the panelists who chose “Unfavorable” as supporting Biden, i.e.
+not supporting Trump.
 
@@
 @fig-vote-biden shows the distribution of vote count in our survey data.
-We see that the survey panelists are heavily in favor of Joe Biden with a over twice of the panelists not favoring Trump.
+We see that the survey panelists are heavily in favor of Joe Biden with over twice the panelists not favoring Trump.
 The bias can be a potential weakness of the dataset and we aim to uncover the factors influencing these political preferences.
@@
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6, margin = margin(0, 10, 0, 10)))
 ```
@@
 Some striking trends we notice in our survey data are more male panelists supporting Trump, majority of Trump supporters being in the 45-59 age group and Trump support declining with increase in education levels.
-While most states seem to support Trump, we see that some states support Trump over Biden.
+While most states seem to support Biden, we see that some states support Trump over Biden.
 This is a fresh change to see over all other factors showing a clear preference for Biden in all categories.
+
+\newpage
 
@@
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6, margin = margin(0, 10, 0, 10)))
 
@@
 For example, we see that females are underrepresented in the survey with the survey having about 49% females whereas the census shows 51% females.
-We also see how population with highest education attained being high school are under-represented whereas population with highest education attained being more than high school are over-representated.
+We also see how population with highest education attained being high school are under-represented whereas population with highest education attained being more than high school are over-represented.
 Difference in proportions like these can be seen in all the variables of interests and this is why post-stratifying is important to ensure precision and better representation of the population.
+
+\newpage
 
@@
 \begin{equation}
-\log\left(\frac{\hat{p}}{1 - \hat{p}}\right) = \beta_0 + \beta_1 \times \text{state} + \beta_2 \times \text{sex} + \beta_3 \times \text{age} + \beta_4 \times \text{education} + \beta_5 \times \text{income} \label{eq:model1}
+\log\left(\frac{\hat{p}}{1 - \hat{p}}\right) = \beta_0 + \beta_1 \times \text{state} + \beta_2 \times \text{sex} + \beta_3 \times \text{age} + \beta_4 \times \text{education} + \beta_5 \times \text{income}
 \end{equation}
@@
   theme(legend.position = "bottom") +
-  theme(axis.text.y = element_text(size = 4, margin = margin(0, 10, 0, 10))) # Adjust size of y-axis text
+  theme(axis.text.y = element_text(size = 5, margin = margin(0, 10, 0, 10)))
 
@@
 and the states on the right are voting for Biden. We see that the majority of the votes for the democratic party
-are found from the Midwest  and the Southern states while only Rhode Island, Oklahoma, Missouri and Nevada are inclined to vote for Trump.
-Rhode Island is almost 75% more likely to vote for Trump, but in competition most of of Alaska, Hawaii, North Dakota, New Mexico and Kansas are inclined to vote for Biden.
+are found from the Midwest and the southern states while only Rhode Island, Oklahoma, Missouri and Nevada are inclined to vote for Trump.
+Rhode Island is almost 75% more likely to vote for Trump, but in competition, most of Alaska, Hawaii, North Dakota, New Mexico and Kansas are inclined to vote for Biden.
 
@@
 
-In @fig-support-age, we used the data to find the proportion of the voters categorized by age, who would vote for the Democratic party.
+In @fig-support-age, we see the proportion of the voters who would vote for the Democratic party categorized by age.
 We see that the younger generation are more likely to vote for Biden compared to the older.
@@
 ```
-@fig-support-sex shows the proportion of Female voters that would vote for the Democratic Party as compared to the Male voters.
+@fig-support-sex shows the proportion of female voters that would vote for the Democratic Party as compared to the male voters.
 We see a 10% difference in the proportion where women are 10% more likely to vote for Biden.
@@
 @fig-support-general looks at the division of voters for each party as an overall.
-We see that even after applying post-stratification, the preference has only changed by a negligible amount with a large majority of the population likely to vote for Trump.
-This graph clearly shows the conclusion we can deduce from our research that it is clear that the population of the United States is more inclined towards the Democratic Party.
+We see that even after applying post-stratification, the preference has only changed by a negligible amount with a large majority of the population likely to vote for Biden.
+This graph clearly shows the conclusion we can deduce from our research, the population of the United States is more inclined towards the Democratic Party.
 
@@
 other than Rhode Island, Missouri, Oklahoma and Nevada are more inclined to vote for the Democratic Party.
-We see that as we go towards the Midwest and Southern Regions, the support for Biden increases as more than 50% of the state is likely to vote for them.
+We see that as we go towards the Midwest and southern regions, the support for Biden increases as more than 50% of the state is likely to vote for them.
 This is similar to the findings in [@James_2016] that address the same prediction for 2024.
@@
 These areas frequently support gun rights, minimal government intrusion, and religious liberty—issues that are strongly aligned with the agenda of the Republican Party, which Trump advocated.
-In contrast, the Northeast and coastline states, which are renowned for their progressive views on culture, urban facilities, and diverse populations, have historically
-tended toward more liberal policies and social ideals that are more in line with the Democratic Party platform, which Biden advocated.
+In contrast, the Northeast and coastline states, which are renowned for their progressive views on culture, urban facilities, and diverse populations, have historically tended toward more liberal policies and social ideals that are more in line with the Democratic Party platform, which Biden advocated.
 
@@
 
-The middle-aged generation i.e individuals of age group from 30-44 years old were inclined to voting for Biden as well.
-Middle-aged voters may align more closely with Biden's policy proposals on issues such as healthcare, economic recovery, climate change, and social justice [@Igielnik_2021] .
+The middle-aged generation i.e.
+individuals of age group from 30-44 years old were inclined to voting for Biden as well.
+Middle-aged voters may align more closely with Biden's policy proposals on issues such as healthcare, economic recovery, climate change, and social justice [@Igielnik_2021].
 
@@
 
-## Contributions to Knowledge:
+## Contributions to Knowledge
 
@@
 
-Lastly, it highlights the high support of the Democratic Party led by Biden from the younger generations, females and people of the midwest and southern states.
+Lastly, it highlights the high support of the Democratic Party led by Biden from the younger generations, females, and people of the Midwest and southern states.
+
+## Weaknesses and Limitations
+
+While our analysis provides valuable insights into voter behavior, it is important to acknowledge several limitations.
+
+Firstly, our reliance on survey data from the Ipsos 2023 Poll introduces a sampling bias, as the sample does not fully represent the entire US population.
+
+On analyzing the counts in the dataset, we see that some of the states have a count of 1 or 2 records in the dataset.
+For example, the District of Columbia and Wyoming, both have one record each in the dataset.
+While these states are not very populated so it is a given that the representation will be lesser, having only one representative leads to a very biased preference from that state.
+The political preference of the state will be considered as being what that one person’s preference is.
+
+If we consider Alaska which has 2 votes, both the representatives from there are supportive of Biden leading to the biased conclusion that Alaska is highly supportive of the Democratic Party.
+This is a contradiction to the fact that Alaska is considered a state which is highly of the Republican Party [@Wikipedia_RedBlue].
+
+Additionally, the survey data does not contain any participants from the state of Vermont and Montana.
+While these states are not very populated [@Wikipedia_2024], their opinions are still important to accurately analyze the preference splits throughout the country.
+
+Consequently, these issues could lead to inaccuracies in our predictions and limit the generalizability of our findings to the broader population.
+
+For our poststratification dataset from ACS, there was no dataset available from 2023.
+Our poststratification data relied on the survey that was conducted in 2022, and Americans' viewpoints on their choices for political parties may have changed in a year.
+
+Additionally, the education variable from ACS had a vague description and only provided a certain level of education.
+The levels that were provided in the dataset were ranging from Grade 5 to 5+ years of college.
+This led to our education levels to be not very diverse but restricted to three major categories: Less than High School, High School and More than High School.
+
+Lastly, while we focused on key demographic variables, other influential variables such as political ideology and campaign messaging were not fully accounted for in our model.
+Political parties use various data from the population such as media trends and ongoing cultural differences in the society.
+They also use micro-targeted political ads to voters on social media and online platforms [@Funakoshi_2020].
+This means that our research has the inaccuracy of consideration of incomplete variable sets.
+
+## Future Directions
+
+Moving forward, further research is needed to address the limitations identified in our study and data to advance our understanding of electoral dynamics.
+Future studies could explore the integration of additional data sources, such as social media sentiment analysis and polling data, to enhance the accuracy of election predictions.
+
+Moreover, there are new strategies opted by political parties in every election that are facilitated by new technologies.
+Exploring the role of media exposure, including news sources, social media platforms, and political messaging, in shaping citizens' attitudes, expectations, and priorities is something that should be considered in shaping the opinions of Americans.
 
@@
 Education levels were standardized to facilitate analysis.
-Respondents' education information (ppeducat) was categorized into three groups: "High school", "More than high school" (including "Bachelors degree or higher" and "Some college"), and "Less than high school".
+Respondents' education information (ppeducat) was categorized into three groups: "High school", "More than high school" (including "Bachelor's degree or higher" and "Some college"), and "Less than high school".
 This was done to match the education level separation available in the poststratification dataset.
@@
 
+\newpage
+
 ## Survey dataset
@@
 
-head(survey_data)
+head(survey_data)  %>%
+  kable(digits = 2)
+
 ```
@@
 
-head(poststrat_data)
+head(poststrat_data)  %>%
+  kable(digits = 2)
 ```

### 2024-03-16 scripts/03-model.R
Update 03-model.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Multilinear logistic regression model
+# Purpose: Bayesian logistic regression model
 # Author: Kaavya Kalani, Monica Sainani
 # Date: 11 March 2024
 # Contact: kaavya.kalani@mail.utoronto.ca, monica.sainani@mail.utoronto.ca

### 2024-03-16 paper/paper.qmd
Final changes

@@
 This poll is based on a nationally representative probability sample of 949 adults aged 18 or older with oversamples among Black and Hispanic respondents.
-The poll is a collection of demographic information and political preferences.
+The poll is a collection of demographic information and political preferences [@citeIpsos].
 
@@
 
-## Post-stratification Data
+## Post-Stratification Data
 
-Our post-stratification data is from the most recent American Community Survey (ACS), the 2022 Sample.
+Our post-stratification data is from the most recent annual American Community Survey (ACS), the 2022 sample, available through IPUMS USA [@citeACS].
 We use this to represent the voter population and adjust the weight of the survey data.
@@
 After we created the model, we then post-stratify these proportions and to conclude who we predict to win the 2024 US Presidential Elections.
-Post-stratification is a statistical technique employed to refine survey estimates by aligning them with known population characteristics.
+Post-Stratification is a statistical technique employed to refine survey estimates by aligning them with known population characteristics.
 It helps us ensure that our findings accurately represent the diversity of the entire population.
@@
   geom_col(fill = "purple") +
-  labs(title = "Proportions of Votes for Biden by age group",
-       y = "Proportion of Votes",
+  labs(y = "Proportion of Votes",
        x = "Age groups")
@@
   geom_col(fill = "red") +
-  labs(title = "Proportions of Votes for Biden by sex",
-       y = "Proportion of Votes",
+  labs(y = "Proportion of Votes",
        x = "Sex groups")
@@
   geom_col(fill = "blue") +
-  labs(title = "Proportions of Votes",
-       y = "Proportion",
+  labs(y = "Proportion",
        x = "Vote") +
@@
 
-For our poststratification dataset from ACS, there was no dataset available from 2023.
-Our poststratification data relied on the survey that was conducted in 2022, and Americans' viewpoints on their choices for political parties may have changed in a year.
+For our post-stratification dataset from ACS, there was no dataset available from 2023.
+Our post-stratification data relied on the survey that was conducted in 2022, and Americans' viewpoints on their choices for political parties may have changed in a year.
 
@@
 Respondents' education information (ppeducat) was categorized into three groups: "High school", "More than high school" (including "Bachelor's degree or higher" and "Some college"), and "Less than high school".
-This was done to match the education level separation available in the poststratification dataset.
+This was done to match the education level separation available in the post-stratification dataset.
 The final column was named education.
@@
 Gender information (ppgender) was just had a change of turning the value to lower case.
-The final column was named sex as just having these two criterias implies that and to maintain with the poststratification data reporting sex as well.
+The final column was named sex as just having these two criterias implies that and to maintain with the post-stratification data reporting sex as well.
 
@@
 
-For the poststratification dataset, the cleaning steps we took were:
+For the post-stratification dataset, the cleaning steps we took were:
 
@@
 
-7. The final poststratification dataset contained the above mentioned 5 columns: age, education, income, state and sex.
+7. The final post-stratification dataset contained the above mentioned 5 columns: age, education, income, state and sex.
 
@@
 ## Survey dataset
-Here is a glimpse of the survey data set used
+Here is a glimpse of the survey dataset used
 
@@
 #| message: false
-
+#| label: tbl-survey
+#| tbl-cap: Cleaned survey dataset
+#| 
 head(survey_data)  %>%
@@
 
-## Poststratification dataset
-Here is a glimpse of the poststratification data set used
+## Post-Stratification dataset
+Here is a glimpse of the post-stratification data set used
 ```{r}
@@
 #| message: false
+#| label: tbl-post-stratification
+#| tbl-cap: Cleaned post-stratification dataset
 
@@
 
+\newpage
+
 # References
